### 2021-07-14 analysis/update_meeting_14_07_2021.Rmd
update meeting

---
title: "Update Meeting: 14-07-2021"
author: "jennysjaarda"
date: "2021-06-14"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
  toc: true
---

```{r setup, include = FALSE}

options(scipen=999)
source("code/settings.R")
load("summary.Rdata")
library(DT)
library(knitr)
library(targets)

```

```{r, echo=FALSE,include = FALSE}
# You need this code to conduct the magic dependences attaching...
DT::datatable(matrix())
```

## Last week summary.

- Update phenotypes, make sure blood chemistry values are included (see list of biomarkers included [here](https://biobank.ndph.ox.ac.uk/ukb/label.cgi?id=17518)).
- Non-urgent analysis: cross partner heritability. 
  - For this we need a full GWAS of index genotype on partner trait (up until this point only run on IVs of interest).
- GWAS of PCs 1-40, adjusted for age and sex, to get IVs (Neale has not run this).
- Run MR in same individual (i.e. standard MR).

## Update.

### Reprocessed PHESANT

- For all new phenotype files since starting the project and added them to the pipeline.

```{r new_phenos, echo = FALSE}

tar_load(data_UKBB_directory)
new_files <- c("ukb39955", "ukb42673", "ukb44073", "ukb44073")

data_UKBB_directory_new <- data_UKBB_directory %>% dplyr::filter(file %in% !!new_files)

DT::datatable(data_UKBB_directory_new, width = "100%")

```

### Identified a problem with the Neale summary stats.
- Only affecting the `low_confidence_variant` column. 
- This filter was applied before generating IVs for each summary file that had been downloaded.
- See a description of the issue on [the Neale github issues page](https://github.com/Nealelab/UK_Biobank_GWAS/issues/20).
- Currently in the process of updating the impacted files:
  - Updated Neale GWAS files can be found here: `/data/sgg2/jenny/data/Neale_UKBB_GWAS/neale_files/variant_flag_update`.  
  - Also will recreate the IVs for these files (~70 of our files), under the same name. 
-  As far as I can tell the issue is only in one direction, where some SNPs that were not really low confidence variants (by their definition) were listed as `low_confidence_variant = TRUE`, not the other way around. In other words, the mistake means that we would only be filtering out variants that we should keep.

### Adding PCs to the pipeline.

- Currently running GWAS of first 40 PCs, residualized for `age` and `sex`. Applied inverse normal rank transformation first.  
- Test between partners.

```{r pc_corr, echo = FALSE}

tar_load(PCs_corr)

DT::datatable(PCs_corr, width = "100%")

```

## Next steps.

Define the following:

**Goal**: To use Mendelian Randomization (MR) to estimate a causal effect of ${X}$ on ${Y}$ between household pairs (i.e. *couples*).

Definitions:

- Let ${X}$ and ${Y}$ denote two random variables representing complex traits. 
- Let ${X}_i$ and ${X}_p$ denote ${X}$ in an index case and in their partner. 
- The genotype data of the SNPs to be used as instrumental variables (IVs) is denoted by ${G}$. 
- Let ${G}_i$ and ${G}_p$ denote ${G}$ in an index case and in their partner. 
- To simplify notation we assume that $E(X) = E(Y) = E(G) =  0$ and $Var(X) = Var(Y) = Var(G) = 1$. 
- The effect sizes of the instruments on $X$ is denoted by ${\beta}$. 
- The causal effect is denoted by ${\alpha}$.
- Let us assume the following models.

$$
\begin{eqnarray*}
	(1) \; X_{i} &=& G_{i}\cdot {\beta_{x, i\rightarrow i} } + \epsilon \\
	(2) \; X_{p} &=& G_{i}\cdot {\beta_{x, i\rightarrow p} } + \epsilon \\
	(3) \; Y_{i} &=& G_{i}\cdot {\beta_{y, i\rightarrow i} } + \epsilon \\
	(4) \; Y_{p} &=& G_{i}\cdot {\beta_{y, i\rightarrow p} } + \epsilon \\
	(5) \; Y_{i} &=& \alpha_j\cdot X_{i} +\epsilon \\
	(6) \; X_{p} &=& \alpha_{x, am}\cdot X_{i{}} +\epsilon \\
	(7) \; Y_{p} &=& \alpha_{y, am}\cdot Y_{i{}} +\epsilon \\
	(8) \; Y_{p} &=& \alpha_{k}\cdot X_{i} +\epsilon

\end{eqnarray*}
$$

- Equation (1) and (3) are simply same-person GWAS result (obtained from Neale).
- Equation (2) and (4) are household GWAS results (calculated in-house).
- Equation (5), and ${\alpha_{j}}$, in the causal effect of a trait ${X}$ on ${Y}$ from a standard MR single-person MR.
- Equation (6) and (7), represent assortative mating effects for traits ${X}$ and ${Y}$, respectively.
- Equation (8) ${\alpha_{k}}$ is the result of phenotype ${X}$ in the index, having a causal effect on a phenotype ${Y}$ in the partner. 
- This effect is either due to the product of one of the two:
  - ${\alpha_{j}}$ with ${\alpha_{x, am}}$ 
  - ${\alpha_{j}}$ with ${\alpha_{y, am}}$