### 2021-07-22 analysis/index.Rmd
Updating website

---
title: "Home"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---

Welcome to my research website.

Check-out my latex tests [here](latex_test.html)