### 2021-08-24 analysis/update_meeting_25_08_2021.Rmd
intial update meeeting 25-08-21

---
title: "Update Meeting: 25-08-2021"
author: "jennysjaarda"
date: "2021-08-24"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
  toc: true
---

```{r setup, include = FALSE}

options(scipen=999)
source("code/settings.R")
library(targets)
library(dplyr)
library(DT)
library(knitr)

```

## Last meeting summary.

## Model summary

```{r proxyMR_DAG, echo = FALSE}
library(knitr)
knitr::include_graphics("assets/proxyMR_DAG.png", error = FALSE)
```

## Results

### Trends amongst bins

```{r MR_binned_het, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
tar_load(household_MR_binned_het)
DT::datatable(household_MR_binned_het, width = "100%")
```

### Standard MR

```{r standard_MR, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
tar_load(standard_MR_summary_BF_sig)
DT::datatable(standard_MR_summary_BF_sig, width = "100%")
```

### Paths from X-index to Y-partner

```{r proxyMR_products, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
tar_load(proxyMR_figure_data)
DT::datatable(proxyMR_figure_data, width = "100%")
```

#### Heterogeneity among paths

```{r proxyMR_figure1, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
tar_load(proxyMR_prod_comparison_figure)
proxyMR_prod_comparison_figure
```

#### Sex heterogeneity amongst paths

``````{r proxyMR_figure2, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
tar_load(proxyMR_sex_comparison_figure)
proxyMR_sex_comparison_figure
```

#### Hierarchical clustering