### 2021-08-31 analysis/index.Rmd
clean up some website docs

---
title: "Home"
site: workflowr::wflow_site
output:
  bookdown::html_document2:
    toc: true
editor_options:
  chunk_output_type: console
---

This is the website for the research project "proxyMR".

If you have cloned the project to a local computer this website is rendered in the `docs` subdirectory of the project directory.

If you are using `workflowr` to publish the research website it will also be rendered online to GitHub Pages.

This page acts as a table of contents for the website. There are links to the web pages generated from the analysis notebooks and to the rendered versions of manuscripts/documents/presentations.

------------------------------------------------------------------------

# Overview documents {.unnumbered}

## [About](about.html) {.unnumbered}

This notebook explains the central ideas behind the project.

## [Notes](notes.html) {.unnumbered}

This notebook is for keeping notes of any points that may be useful for
later project or manuscript development and which are not covered in the
analysis notebooks or at risk of getting lost in the notebooks.

## [Workflow management](workflow.html) {.unnumbered}

This project uses the [`targets`](https://wlandau.github.io/targets/) and [`workflowr`](https://github.com/jdblischak/workflowr) packages for managing the workflow of the project (making sure that the dependencies between computational steps are satisfied). This notebook contains notes on the proposed workflow for
using `targets` and `workflowr`.

## [Meetings](meetings.html) {.unnumbered}

## [Project Workflow Status](m_00_status.html) {.unnumbered}

This notebook displays the computational status of the project, that is,
whether everything is up to date. Ironically, this is currently the only
notebook that must be run manually, so this notebook only displays the
status **when the notebook was last executed** and there is no
indication whether the project status has changed since then.

------------------------------------------------------------------------

# Publications and Presentations {.unnumbered}

Links to rendered manuscripts and presentations will go here.

------------------------------------------------------------------------

# Analayses Notebooks {.unnumbered}

These notebooks capture the analyses that were carried out to develop
the code of the core processing pipeline. They are organised as
side-chains to the core processing pipeline.

Typically, a meta notebook will analyse the data available at one stage
of the core pipeline, to guide the writing of the functions required to
get to the next stage of the core pipeline. These meta notebooks
generally conclude with the definition of a function that will be used
in the core pipeline.

There may be multiple notebooks all relating to different aspects of the
same stage of the core pipeline.

Sometimes the analyses are more diffuse - characterising the data in a
way that may be helpful for guiding the development of future core
stages, but not immediately resulting in the development of functions
for the core pipeline.

## 01 Read, check, and standardise the entity data {.unnumbered}

Determine the initial data preparation of the imported entity records.

### [m_01_1\_read_raw_entity_data](m_01_1_read_raw_entity_data.html) {.unnumbered}

Read the raw entity data.

### [m_01_2\_exclusions](m_01_2_exclusions.html) {.unnumbered}

Apply the row exclusions (exclude test records and all records with
status other than ACTIVE & VERIFIED).

### [m_01_3\_drop_novar](m_01_3_drop_novar.html) {.unnumbered}

Drop variables with no variation.

### [m_01_4\_parse_dates](m_01_4_parse_dates.html) {.unnumbered}

Parse the date columns.

### [m_01_5\_check_admin](m_01_5_check_admin.html) {.unnumbered}

Check the administrative variables (data relating to the administration
of voter registration).

### [m_01_6\_check_resid](m_01_6_check_resid.html) {.unnumbered}

Check the residence variables - residential address and phone number.

### [m_01_7\_check_demog](m_01_7_check_demog.html) {.unnumbered}

Check the demographic variables - sex, age, and birth place.

### [m_01_8\_check_name](m_01_8_check_name.html) {.unnumbered}

Check the name variables.

### [m_01_9\_clean_vars](m_01_9_clean_vars.html) {.unnumbered}

Clean all the variables.

------------------------------------------------------------------------

# Acknowledgements

Primary analysis performed by: [Jenny Sjaarda](jennysjaarda@gmail.com).

Project supervisor: [Professor Zoltan Kutalik](https://wp.unil.ch/sgg/people/zoltan/).

This work has been supported by the [Swiss National Science Foundation](http://www.snf.ch/en/funding/Pages/default.aspx).