### 2021-09-06 analysis/update_meeting_06_09_2021.Rmd
update meeting 6/09/21

---
title: "Update Meeting: 06-09-2021"
author: "jennysjaarda"
date: "2021-09-06"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}

options(scipen=999)
source("code/settings.R")
library(targets)
library(dplyr)
library(DT)
library(knitr)

```

# Model summaries

```{r proxyMR_DAG, echo = FALSE}
library(knitr)
knitr::include_graphics("assets/proxyMR_DAG.png", error = FALSE)
knitr::include_graphics("assets/proxyMR_DAG_MV.png", error = FALSE)

```

```{r load_targets, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

tar_load(household_MR_binned_het)
tar_load(num_tests_by_PCs)
tar_load(standard_MR_summary_BF_sig)
tar_load(household_MR_summary_BF_sig)
tar_load(proxyMR_comparison_summary)

```

# Questions

1. Impact of meta-analyzing after MR is calculated?

2. Trend across bins: should it be a weighted linear model?

3. Less stringent BF correction: 
    - Calculate correlation matrix of all 131 traits tested. 
    - Perform PCs on the resulting corr matrix. 
    - Number of tests = # of PCs needed until variance explained is > 99.5%.
    - Resulting # of tests: `r num_tests_by_PCs`.

4. Filter by same-person MR or proxyMR?

5. Summed `rho` and `gamma` and compared this to `omega`, for the variance - it's just the sum of these two variances?

6. MV MR: X -> Z -> Y
    - Drop Z's if not significant in MV MR (i.e. `Y ~ X + Z1 + Z2 + Z3 + ...`)
    - Do we drop from the model itself and recompute the beta-coefficients? Or simply not include them in the result sum?

# AM summary

```{r AM_summary, echo = TRUE, include = TRUE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

AM_result <- household_MR_binned_het %>% filter(same_trait)

AM_sig_sex_het <- AM_result %>% filter(sex_het_pval < 0.05/num_tests_by_PCs) %>% dplyr::select(exposure_ID, exposure_sex, IVW_beta, IVW_se, sex_het_pval) %>% unique()

AM_sig_age_slope <- AM_result %>% filter(grouping_var=="age_even_bins") %>% filter(bin_slope_pval_meta < 0.05/num_tests_by_PCs) %>%
  dplyr::select(exposure_ID, exposure_sex, IVW_beta, IVW_se, bin_slope_beta_meta, bin_slope_pval_meta) %>% unique()

AM_sig_tt_slope <- AM_result %>% filter(grouping_var=="time_together_even_bins") %>% filter(bin_slope_pval_meta < 0.05/num_tests_by_PCs) %>%
  dplyr::select(exposure_ID, exposure_sex, IVW_beta, IVW_se, bin_slope_beta_meta, bin_slope_pval_meta) %>% unique()

AM_sig_age_sex_het <- AM_result %>% filter(grouping_var=="age_even_bins") %>% filter(bin_slope_sex_het_pval < 0.05/num_tests_by_PCs) %>%
  dplyr::select(exposure_ID, exposure_sex, IVW_beta, IVW_se, bin_slope_beta, bin_slope_pval, bin_slope_beta_meta, bin_slope_pval_meta) %>% unique()

AM_sig_tt_sex_het <- AM_result %>% filter(grouping_var=="time_together_even_bins") %>% filter(bin_slope_sex_het_pval < 0.05/num_tests_by_PCs) %>%
  dplyr::select(exposure_ID, exposure_sex, IVW_beta, IVW_se, bin_slope_beta, bin_slope_pval, bin_slope_beta_meta, bin_slope_pval_meta) %>% unique()

```

# ProxyMR

Restricting XY pairs: 
  - If we base it on a significant MR in the same individuals (same MR: $X \rightarrow Y$), the number of tests is: `r dim(standard_MR_summary_BF_sig)[1]`.
  - If we base it on a significant **proxyMR** ($X_{i} \rightarrow Y_{p}$), the number of tests is: `r dim(household_MR_summary_BF_sig)[1]`.

Can we use the PC multiple-test correction we applied above here, or just use conservative raw BF correction?

Comparison of omega to rho + gamma instead of just rho or gamma alone. There are more significant results using both.

```{r proxy_MR_update, echo = TRUE, include = TRUE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

table(proxyMR_comparison_summary$omega_vs_gam_BF_sig_meta)

table(proxyMR_comparison_summary$omega_vs_rho_BF_sig_meta)

table(proxyMR_comparison_summary$omega_vs_gam_rho_BF_sig_meta)

#which(proxyMR_comparison_summary$omega_vs_gam_rho_BF_sig_meta & !proxyMR_comparison_summary$omega_vs_rho_BF_sig_meta)[1:10]

t(proxyMR_comparison_summary[205,c("exposure_ID", "outcome_ID", "exposure_description", "outcome_description", "gam_meta_beta", "rho_meta_beta", "omega_meta_beta", "gam_rho_meta_beta",
                                    "gam_meta_se", "rho_meta_se", "omega_meta_se", "gam_rho_meta_se",
                                   "omega_vs_gam_meta_p", "omega_vs_rho_meta_p", "omega_vs_gam_rho_meta_p")])

```