### 2021-09-14 analysis/notes.Rmd

---
title: "Notes"
author: "jennysjaarda"
date: "2021-09-08"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

# Meta-analysis across sexes

There are two inputs into MR:

1. SNP/exposure effects. 
2. SNP/outcome effects.

If we want to perform a MR between exposure $X$ and outcome $Y$ comprising both sexes, we have three options:

1. Use joint GWAS statistics (i.e. $X \sim G$ is run in sexes combined, adjusted for sex).
2. Use sex-specific GWAS statistics and then meta-analyze **before** running the MR. 
3. Use sex-specific GWAS statistics and then meta-analyze the **after** running the MR.

In other words, there are three options for pooling the data:

1. At the selection of subgroup for GWAS analysis level. 
2. At the SNP/exposure and SNP/outcome level.
3. At the MR level.