### 2021-09-14 analysis/index.Rmd

---
title: "Home"
site: workflowr::wflow_site
output:
  bookdown::html_document2:
    toc: true
editor_options:
  chunk_output_type: console
---

This is the website for the research project "proxyMR".

If you have cloned the project to a local computer this website is rendered in the `docs` subdirectory of the project directory.

If you are using `workflowr` to publish the research website it will also be rendered online to GitHub Pages.

This page acts as a table of contents for the website. There are links to the web pages generated from the analysis notebooks and to the rendered versions of manuscripts/documents/presentations.

------------------------------------------------------------------------

# Overview documents {.unnumbered}

## [About](about.html) {.unnumbered}

This notebook explains the central ideas behind the project.

## [Notes](notes.html) {.unnumbered}

This notebook is for keeping notes of any points that may be useful for
later project or manuscript development.

## [To do](todo_list.html) {.unnumbered}

Current tasks and completed tasks (roughly organized by dates/weeks).

## [Workflow management](workflow.html) {.unnumbered}

This project uses the [`targets`](https://wlandau.github.io/targets/) and [`workflowr`](https://github.com/jdblischak/workflowr) packages for managing the workflow of the project (making sure that the dependencies between computational steps are satisfied). This notebook contains notes on the proposed workflow for
using `targets` and `workflowr`.

## [Meetings](meetings.html) {.unnumbered}

Weekly meeting were held roughly once per week or every other week.

------------------------------------------------------------------------

# Publications and Presentations {.unnumbered}

- Group Meeting [07-09-2021](presentations/group_meeting_070921/group_meeting_proxyMR_070921.pdf)

------------------------------------------------------------------------

# Analyses Notebooks {.unnumbered}

These notebooks capture the analyses that were carried out to develop
the code of the core [processing pipeline](pipeline_overview.html).

## 1. Prepare data {.unnumbered}

Prepare phenotypes and couples for analysis in MR pipeline.

### [Couple selection](couple_selection.html) {.unnumbered}

Identify couples in the UKBB.

### [Phenotype selection](phenotype_selection.html) {.unnumbered}

Select phenotypes for analysis in pipeline.

### [Final phenotype list](final_phenotype_list.html) {.unnumbered}

Final list of phenotypes for analysis in pipeline and some basic characteristics (correlation among couples, number of independent IVs).

## 2. Run MR pipeline {.unnumbered}

### [MR between partners (partner vs index)]() {.unnumbered}

- Identify SNP-exposure effects (GW-significance) for phenotypes (sex specific) from Neale database [$X_{i} \sim G_{i}$].
- Test SNP-outcome effects between genetic instruments and partner outcome phenotypes (sex specific) [$Y_{p} \sim G_{i}$].
- Combine in MR to assess causal relationships between index and partner [$Y_{p} \sim X_{i}$].

### [Assortative mating MR]() {.unnumbered}

- Restrict to only MRs where exposure is the same as outcome (for downstream analyses) [$X_{p} \sim X_{i}$].

### [MR within same individual]() {.unnumbered}

- Identify SNP-exposure effects (GW-significance) for phenotypes (sex specific) from Neale database [$X_{i} \sim G_{i}$].
- Pull SNP-outcome effects  between genetic instruments and outcome phenotypes (sex specific) from Neale database [$Y_{i} \sim G_{i}$].
- Combine in MR to assess causal relationship between all phenotypes [$Y_{i} \sim X_{i}$].

### [Compare paths towards outcome in partner]() {.unnumbered}

- Compare paths from index exposure phenotype to partner outcome phenotype [$X_{i} \rightarrow Y_{p}$].

### [PC correlation among couples]() {.unnumbered}

------------------------------------------------------------------------

# Acknowledgements {.unnumbered}

Primary analysis performed by: [Jenny Sjaarda](jennysjaarda@gmail.com).

Project supervisor: [Professor Zoltan Kutalik](https://wp.unil.ch/sgg/people/zoltan/).

This work has been supported by the [Swiss National Science Foundation](http://www.snf.ch/en/funding/Pages/default.aspx).