### 2021-09-14 analysis/final_phenotype_list.Rmd
update some docs

---
title: "Final phenotype list"
author: "Jenny Sjaarda"
date: "2021-09-02"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}

options(scipen=999)
source("code/settings.R")
library(targets)
library(dplyr)
library(DT)
library(knitr)
library(plotly)
library(ggplot2)

```

```{r load_targets, include = FALSE}

tar_load(traits_final)

```

After [phenotype selection](phenotype_selection.html), `r dim(traits_final)[1]` traits remained for downstream analyses.

The final set of traits and some basic properties of interest can be seen in the table below.

```{r final_phenotypes, echo=FALSE}

traits_final.DT <- traits_final %>% as_tibble() %>% mutate(r = sqrt(as.numeric(r2))) %>% dplyr::select(Neale_pheno_ID, description, variable_type, category, N_pairs, r2, r, num_IVs, num_IVs_pass_het)
DT::datatable(traits_final.DT, width = "100%") %>% formatRound(columns=c('r2', 'r'), digits = 3)

```

------------------------------------------------------------------------

A few summaries of various properties are show below.

```{r final_phenotypes_summaries, echo = FALSE}
variable_type_tab <- traits_final.DT %>% count(variable_type)
category_tab <- traits_final.DT %>% count(category)

knitr::kable(variable_type_tab, format = "pandoc")
knitr::kable(category_tab, format = "pandoc")

```

------------------------------------------------------------------------

```{r IV_freq, echo = FALSE}

ggplot(traits_final.DT, aes(as.numeric(num_IVs_pass_het))) + geom_histogram(binwidth = 10, alpha=0.5) + theme_classic()  + xlab("Number of IVs") + ylab("Frequency")

ggplot(traits_final.DT, aes(as.numeric(r))) + geom_histogram(binwidth = 0.01, alpha=0.5) + theme_classic()  + xlab("Correlation (r)") + ylab("Frequency")

```