### 2021-09-14 analysis/phenotype_selection.Rmd

---
title: "Phenotype selection"
author: "Jenny Sjaarda"
date: "2021-09-02"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}

options(scipen=999)
source("code/settings.R")
library(targets)
library(dplyr)
library(DT)
library(knitr)

```

```{r load_targets, include = FALSE}

tar_load(data_phesant_directory)
tar_load(Neale_SGG_dir_filt)
tar_load(traits_corr)
tar_load(traits_corr2)
tar_load(traits_corr3)
tar_load(traits_corr4)
tar_load(traits_corr5)

```

**Phenotypes were selected according to the following scheme**

*Note that as none of the filtering criteria apply any BF correction for the number of phenotypes, the order of the steps does not matter.*

# UKBB Phenotypes in internal database processed by PHESANT.

All SGG phenotypes went internal processing according the PHESANT pipeline. Some traits were excluded for various reasons (genetic trait, insufficient sample size, etc.).

Number of traits initially = **`r length(unique(data_phesant_directory[,2]))`**.

Traits included and their correlations amongst couples can be seen below.

```{r initial_corr, echo=FALSE}

traits_corr.DT <- traits_corr %>% as_tibble() %>% mutate_all(parse_guess) %>% mutate(r = sqrt(r2))
DT::datatable(traits_corr.DT, width = "100%") %>% formatRound(columns=c('r2', 'r'), digits = 3)
```

# Merge with Neale summary stats.

Limited to only traits that were processed by Neale and had both sex-specific and joint summary statistics available (most traits that were excluded did not have all 3 subgroups were not available, some other traits were excluded by the Neale group).

Number remaining after filter = **`r length(unique(Neale_SGG_dir_filt$phenotype))`**.

The traits that *are* in the SGG database but were **not** included due to Neale filtering can be seen below. Can compare with the [Neale Manifest](https://docs.google.com/spreadsheets/d/1kvPoupSzsSFBNSztMzl04xMoSC3Kcx3CrjVf4yBmESU/edit?ts=5b5f17db#gid=178908679) (search by ID) to get a sense of why traits were excluded (likely there is not a GWAS file for `both_sexes`, `male` and `female`).

```{r SGG_Neale_non_link, echo=FALSE}

## traits corr is the same dimension as (1) above
Neale_SGG_dir_filt_unique <- Neale_SGG_dir_filt[,c("phenotype", "phenotype_sub", "sgg_phesant_name", "description", "SGG_available", "variable_type")] %>% unique()
non_merged <- anti_join(traits_corr, Neale_SGG_dir_filt_unique, by = c("ID"="sgg_phesant_name")) %>% as_tibble() %>% mutate_all(parse_guess) %>% mutate(r = sqrt(r2))
DT::datatable(non_merged, width = "100%") %>% formatRound(columns=c('r2', 'r'), digits = 3)

```

# Evidence of correlation among couples.

Traits were excluded if the correlation among couples was < `r household_correlation_threshold` (i.e. $\sqrt{(r^2)}$).

Number remaining after filter = **`r dim(traits_corr2)[1]/3`**.

The traits remaining after this filter can be seen below.

```{r corr_filter, echo=FALSE}

traits_corr2.DT <- traits_corr2 %>% as_tibble() %>% mutate(r = sqrt(r2)) %>% dplyr::select(Neale_pheno_ID, description, variable_type, category, N_pairs, r2, r) %>% unique()
DT::datatable(traits_corr2.DT, width = "100%") %>% formatRound(columns=c('r2', 'r'), digits = 3)
```

# Sufficient power for MR based on number of IVs.

Phenotypes were excluded if the number of valid IV was < `r num_IVs_threshold`. IVs were defined based on a p-value of p < `r IV_threshold` in `both_sexes` after clumping (performed in PLINK with the options `--clump-kb 10000` and `--clump-r2 0.001` using the 1000 Genomes European samples as a reference).

Number remaining after filter = **`r dim(traits_corr3$to_run)[1]`**.

The traits remaining after IV filtering can be seen below.

```{r IV_filter, echo=FALSE}

traits_corr3_filt <- traits_corr3$to_run 
traits_corr3_filt.DT <- traits_corr3_filt %>% as_tibble() %>% mutate(r = sqrt(as.numeric(r2))) %>% dplyr::select(Neale_pheno_ID, description, variable_type, category, N_pairs, r2, r, num_IVs)
DT::datatable(traits_corr3_filt.DT, width = "100%") %>% formatRound(columns=c('r2', 'r'), digits = 3)
```

# No evidence of sex-heterogeneity among valid IVs.

Using the sex-specific summary statistics, heterogeneity between sexes was calculated. IVs that showed significant evidence of heterogeneity between sexes were included (p < 0.05/[number of IVs]). After filtering IVs with significant heterogeneity, phenotypes were filtered to again have IVs < `r num_IVs_threshold`.

Number remaining after filter = **`r dim(traits_corr4$to_run)[1]`**.

The traits remaining after sex-heterogeneity filtering can be seen below.

```{r sex_het_filter, echo=FALSE}

traits_corr4_filt <- traits_corr4$to_run 
traits_corr4_filt.DT <- traits_corr4_filt %>% as_tibble() %>% mutate(r = sqrt(as.numeric(r2))) %>% dplyr::select(Neale_pheno_ID, description, variable_type, category, N_pairs, r2, r, num_IVs, num_IVs_pass_het)
DT::datatable(traits_corr4_filt.DT, width = "100%") %>% formatRound(columns=c('r2', 'r'), digits = 3)
```

# Non-binary.

Phenotypes were limited to `continuous_irnt` or `ordinal` (i.e. `binary` traits were removed).

Number remaining after filter = **`r dim(traits_corr5)[1]`**.

The traits remaining after sex-heterogeneity filtering can be seen below.

```{r continuous_filter, echo=FALSE}

traits_corr5.DT <- traits_corr5 %>% as_tibble() %>% mutate(r = sqrt(as.numeric(r2))) %>% dplyr::select(Neale_pheno_ID, description, variable_type, category, N_pairs, r2, r, num_IVs, num_IVs_pass_het)
DT::datatable(traits_corr5.DT, width = "100%") %>% formatRound(columns=c('r2', 'r'), digits = 3)

```