### 2021-09-14 analysis/todo_list.Rmd

---
title: "To do"
author: "jennysjaarda"
date: "2021-09-08"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

# 06/09/2021

- ~~Sum $\rho$ and $\gamma$.~~
- ~~Meta-analyzed binned results across sexes.~~ 
- ~~Fix variance calculation mistake in calculating $\rho$ and $\gamma$.~~
- ~~Write script/function which replaces `Neale_ID` with trait description.~~
- ~~Clean up `index.Rmd`~~
- ~~Clean up `about.Rmd`~~
- Add Rmd building to `targets` pipeline. 
- MVMR. 
- AM MR summary Rmd.
- PC correlation among couples summary Rmd. 
- Clean up update meeting Rmd.

# 08/09/2021

- ~~Weight trend line by inverse variance.~~ 
- Update AF in Neale summary stats. 
- Create couple selection Rmd. 
- Number of tests for filtering by $\omega$ should exclude those with only one trait. (i.e. number of tests would be $j^2 - j$).
- ~~Filter by $\omega$ instead of same-person MR.~~ 
- Force intercept through 0 in omega vs gamma/rho plot. 
- Explore impact of meta-anlaysis at SNP/trait level vs. MR-level.
  - Meta-analyze standard MR SNP/trait estimates across sexes. 
- Hierarchy figure (add labels).
- Make visio figures in LucidChart.
- Explore IV overlap between X and Y:
    - Determine percentage of IV overlap between X IVs and Y IVs. 
    - MV MR, add SNPs/trait effect from $X_{i} \rightarrow Y_{p}$ to $Y_{i} \rightarrow Y_{p}$ MR (i.e. run a MV MR of the form: $Y_{p} \sim Y_{i} + X_{i}$).
- Add script/function which replaces `Neale_ID` with trait description to main functions.