### 2021-09-15 analysis/geographical_impact.Rmd
adding Rmd of geographical impact

---
title: Assessing impact of geography as confounder in causal relationships between couples
author: "jennysjaarda"
date: "2021-09-14"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}

options(scipen=999)
source("code/settings.R")

library(DT)
library(knitr)
library(targets)
library(tidyverse)
```

```{r load_targets, include = FALSE}

tar_load(data_id_sex)
tar_load(PCs_corr)

```

# Background

- Geographical location is clearly a strong indicator of couple choice. 
- Not only do couples tend to choose partners that are phenotypically similar to them (because of common interests, political views, etc.), but also because of location (i.e. convenience).
- Geographical similarity induces genetic and in turn phenotypic similarity. 
- We sought to explore the impact of genography on mate-choice by adding PCs to the pipeline.

# Analysis

## GWAS of PCs

- GWAS of first 40 PCs run using [bgenie](https://jmarchini.org/bgenie/).
- First, inverse normal rank transformation (INT) was applied within genetic, consenting, European sample (N = `r dim(data_id_sex)[1]`) before performing GWAS (this was done before dividing into sex-specific subgroups).
- After INT, PCs were residualized for `age` and `sex`.
- Run in:
  - Entire sample (residualized for both age and sex).  
  - In sex-specific subgroups (residualized for only age). 
- This was run in the entire genetic, consenting, European UKB sample (not just couples).
- GWAS was **not** adjusted for PCs (as is typical in GWAS) because PCs are linearly independent of each other (although the sample used for PC calculation - by the UKB investigators - would be slightly different that our GWAS sample, so the PCs would not necessarily be completely linearly independent in this sample).

## Correlation of PCs between couples

- Test raw "phenotypic" correlation of PCs between partners in UKB. 
- As above, PCs were transformed using INT within genetic, consenting, European sample (N = `r dim(data_id_sex)[1]`) and correlations were calculated within couples on the transformed PCs.

```{r pc_corr, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

PCs_corr.DT <- PCs_corr %>% mutate(r = sqrt(as.numeric(r2))) 
DT::datatable(PCs_corr, width = "100%")  %>% formatSignif(columns=c('r2', 'r'), digits=3)

```