### 2021-09-15 analysis/update_meeting_25_08_2021.Rmd

---
title: "Update Meeting: 25-08-2021"
author: "Jenny Sjaarda"
date: "2021-08-24"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
  toc: true
---

```{r setup, include = FALSE}

options(scipen=999)
source("code/settings.R")
source("R/functions.R")
library(targets)
library(dplyr)
library(DT)
library(knitr)

```

```{r load_targets, include = FALSE}

tar_load(household_MR_binned_het)
tar_load(traits_final)
tar_load(standard_MR_summary_BF_sig)
tar_load(traits_corr2_filled)
tar_load(proxyMR_comparison_summary)
tar_load(proxyMR_prod_comparison_figure)
tar_load(proxyMR_sex_comparison_figure)
```

# Last meeting summary.

- Discussed summary of pathways from $X$ in index ($X_i$) to $Y$ in the partner ($Y_p$).
- Need for running the same individual MR (i.e. simply the standard MR) so we can compare the two pathways (see DAG below).
- Multiply perpendicular paths together and to calculate the variance between them use the formula [here]( https://stats.stackexchange.com/questions/397839/varxy-if-x-and-y-are-independent-random-variables).

# Model summary.

```{r proxyMR_DAG, echo = FALSE}
library(knitr)
knitr::include_graphics("assets/two_trait_model.png", error = FALSE)
```

# Results.

## Heterogeneity amonst different groups.

### Sex heterogeneity

For each proxyMR ($Y_{p} \sim X_i$), analyses were run in each sex separately. The results below display the results that differ among sexes, after BF correction (i.e. `0.05 / (`r dim(traits_final)[1]` * `r dim(traits_final)[1]`)`)

```{r sig_sex_het, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}

household_MR_binned_het_temp <- replace_Neale_ID(household_MR_binned_het, traits_corr2_filled, "exposure_ID", "exposure_description")
household_MR_binned_het_temp2 <- replace_Neale_ID(household_MR_binned_het_temp, traits_corr2_filled, "outcome_ID", "outcome_description")
household_MR_binned_het <- household_MR_binned_het_temp2

num_tests <- dim(traits_final)[1]^2
sex_het_BF_result <- household_MR_binned_het %>% filter(grouping_var=="age_even_bins") %>% 
  dplyr::select(exposure_description, outcome_description, exposure_sex, outcome_sex, same_trait, IVW_beta, IVW_pval, sex_het_pval) %>% filter(as.numeric(sex_het_pval) < 0.05/num_tests)
DT::datatable(sex_het_BF_result, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval', 'sex_het_pval'), digits=3)
```

However in this case, it may be more sensible to only consider same traits (i.e. testing for more of an "assortative mating" effect), reducing the number of tests to ``r dim(traits_final)[1]``. However, none pass BF-significance (p < `0.05/`r dim(traits_final)[1]` = `r 0.05/dim(traits_final)[1]``). The most significant results is `lifetime number of sexual partners`.

```{r sex_het_same_trait, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}

sex_het_same_trait_result <- household_MR_binned_het %>% filter(grouping_var=="age_even_bins") %>% filter(same_trait==TRUE) %>%
 dplyr::select(exposure_description, exposure_sex, outcome_sex, same_trait, IVW_beta, IVW_pval, sex_het_pval) %>% arrange(sex_het_pval)

DT::datatable(sex_het_same_trait_result, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval', 'sex_het_pval'), digits=3)
```

### Binned heterogeneity

For each proxyMR ($Y_{p} \sim X_i$), MR analyses were run in the full sample as well as in 5 roughly equal sized bins according to `time_together` and `age`. Specifically, the outcome data set was split into 5 bins, and the SNP-outcome effect was estimated in each bin separately. These outcome-SNP effects were then used to generate bin-specific MR estimates, using the same SNP-exposure effects from Neale.

To estimate the difference among bins, used two approaches:

  1. The Cochran's Q test to test for heterogeneity in meta-analyses.
  2. Tested the significance of the slope of linear model of median bin (either `age` or `time together`) versus the bin-specific MR estimate.

The two results are shown below, restricted to only same_traits. Note that each statistic was computed in each sex separately and meta-analyzed.

#### Q-statistic results

```{r bin_q_het, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}

q_het_result_age <- household_MR_binned_het %>% filter(same_trait==TRUE) %>% filter(grouping_var=="age_even_bins") %>% 
  dplyr::select(exposure_description, exposure_sex, outcome_sex, grouping_var, IVW_beta, IVW_pval, Q_pval, Q_pval_meta)

q_het_result_tt <- household_MR_binned_het %>% filter(same_trait==TRUE) %>% filter(grouping_var=="time_together_even_bins") %>% 
  dplyr::select(exposure_description, exposure_sex, outcome_sex, grouping_var, IVW_beta, IVW_pval, Q_pval, Q_pval_meta)

DT::datatable(q_het_result_age, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval', 'Q_pval', 'Q_pval_meta'), digits=3)
DT::datatable(q_het_result_tt, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval', 'Q_pval', 'Q_pval_meta'), digits=3)
```

#### Slope results

```{r bin_slope_het, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}

slope_het_result_age <- household_MR_binned_het %>% filter(same_trait==TRUE) %>% filter(grouping_var=="age_even_bins") %>% 
  dplyr::select(exposure_description, exposure_sex, outcome_sex, grouping_var, IVW_beta, IVW_pval, bin_slope_beta, bin_slope_pval, bin_slope_meta_beta, bin_slope_meta_pval, bin_slope_sex_het_pval)

slope_het_result_tt <- household_MR_binned_het %>% filter(same_trait==TRUE) %>% filter(grouping_var=="time_together_even_bins") %>% 
  dplyr::select(exposure_description, exposure_sex, outcome_sex, grouping_var, IVW_beta, IVW_pval, bin_slope_beta, bin_slope_pval, bin_slope_meta_beta, bin_slope_meta_pval, bin_slope_sex_het_pval)

DT::datatable(slope_het_result_age, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval', 'bin_slope_beta', 'bin_slope_pval', 'bin_slope_meta_beta', 'bin_slope_meta_pval', 'bin_slope_sex_het_pval'), digits=3)
DT::datatable(slope_het_result_tt, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval', 'bin_slope_beta', 'bin_slope_pval', 'bin_slope_meta_beta', 'bin_slope_meta_pval', 'bin_slope_sex_het_pval'), digits=3)
```

## Standard MR

To complete all MR anlayses in the DAG above, needed to generate same-person MR estimates (i.e. standard MR). These were also run sex-specific, using Neale summary statistics. As expected, there are many BF-significant summary MR results. Significance was determined based on the meta-analyzed result across sexes. Thus the number of tests was equal to: ``r dim(traits_final)[1]`^2 - `r dim(traits_final)[1]` = `r dim(traits_final)[1]^2 - dim(traits_final)[1]` (number of traits = ``r dim(traits_final)[1]``). The BF-significant results can be seen below.

```{r standard_MR, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

standard_MR_summary_BF_sig_sub <- standard_MR_summary_BF_sig %>% dplyr::select(exposure_description, outcome_description, N_snps, IVW_meta_beta, IVW_meta_pval) %>% unique()
DT::datatable(standard_MR_summary_BF_sig_sub, width = "100%") %>% formatSignif(columns=c('IVW_meta_beta', 'IVW_meta_pval'), digits=3)
```

Number of BF-signifcant MR pairs: `r dim(standard_MR_summary_BF_sig)[1]`.

## Paths from X-index to Y-partner

There are three distinct paths from $X_{index}$ to $Y_{partner}$ as illustrated in the DAG above:

  1. Assortative mating through $X$: $\gamma$
  2. Assortative mating through $Y$: $\rho$
  3. Indirectly (i.e. through a latent variable which either increases or decreases the observed effect of $X_{index}$ to $Y_{partner}$): $\omega$

The estimates of $\gamma$, $\rho$ and $\omega$ can be seen below, meta-analyzed across sex.

```{r proxyMR_products, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), eval = TRUE}

proxyMR_comparison_summary.DT <- proxyMR_comparison_summary %>% filter(exposure_sex=="male") %>% 
  dplyr::select(exposure_ID, outcome_ID, exposure_description, outcome_description, 
                gam_meta_beta, gam_meta_pval, gam_sex_het_pval, 
                rho_meta_beta, rho_meta_pval, rho_sex_het_pval,
                rho_meta_beta, rho_meta_pval, rho_sex_het_pval)

DT::datatable(proxyMR_comparison_summary.DT, width = "100%") %>% formatSignif(columns=c('gam_meta_beta', 'gam_meta_pval', 'gam_sex_het_pval', 
                                                                                        'rho_meta_beta', 'rho_meta_pval', 'rho_sex_het_pval',
                                                                                        'rho_meta_beta', 'rho_meta_pval', 'rho_sex_het_pval'), digits=3)
```

### Heterogeneity among paths

Next, we wanted to compare the three paths from $X_index$ to $Y_partner$. This was done using a standard heterogeneity test.

```{r proxyMR_product_comparison, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), eval = TRUE}

proxyMR_comparison_summary_prod <- proxyMR_comparison_summary %>%
  dplyr::select(exposure_description, outcome_description, 
                exposure_category, outcome_category, exposure_sex,
                omega_vs_gam_p, omega_vs_gam_meta_p,
                omega_vs_rho_p, omega_vs_rho_meta_p,
                gam_vs_rho_p, gam_vs_rho_meta_p,
                omega_vs_gam_rho_p, omega_vs_gam_rho_meta_p)

DT::datatable(proxyMR_comparison_summary_prod, width = "100%") %>% formatSignif(
  columns = c(
    'omega_vs_gam_p',
    'omega_vs_gam_meta_p',
    'omega_vs_rho_p',
    'omega_vs_rho_meta_p',
    'gam_vs_rho_p',
    'gam_vs_rho_meta_p',
    'omega_vs_gam_rho_p',
    'omega_vs_gam_rho_meta_p'
  ),
  digits = 3
)
```

``````{r proxyMR_figure1, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600}
proxyMR_prod_comparison_figure
```

### Sex heterogeneity amongst paths

``````{r proxyMR_figure2, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600}
proxyMR_sex_comparison_figure
```

### Hierarchical clustering

Working on a figure like the one [here](https://www.r-graph-gallery.com/310-custom-hierarchical-edge-bundling.html) and [here](https://www.r-graph-gallery.com/311-add-labels-to-hierarchical-edge-bundling.html).