### 2021-09-24 analysis/update_meeting_23_09_2021.Rmd
add update meeting 23-09-2021

---
title: "Update Meeting: 23-09-2021"
author: "Jenny Sjaarda"
date: "2021-09-20"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}

options(scipen=999)
source("code/settings.R")
library(targets)
library(dplyr)
library(DT)
library(knitr)
library(ggplot2)

```

```{r load_targets, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

tar_load(proxyyMR_IV_overlap)
tar_load(proxyMR_MR_paths_summary_yiyp_adj)
tar_load(proxyMR_comparison_summary_yiyp_adj)
```

# Last meeting summary.

- Because the results comparing $\omega$ to $\rho + \gamma$ were slightly confusing (see last meeting summary [here](https://jennysjaarda.github.io/proxyMR/update_meeting_06_09_2021.html)), decided we should explore the impact of $X_i$ on the $Y_i \rightarrow Y_p$ causal effects.

```{r two_trait_model, echo = FALSE}
library(knitr)
knitr::include_graphics("assets/two_trait_model.png", error = FALSE)
```

- Two tasks to explore:
    1. Identify overlap between $X_i \rightarrow X_p$ IVs and $Y_i \rightarrow Y_p$. 
    2. Adjust the $Y_i \rightarrow Y_p$ MR results for effects of $X_i$ (as depicted below).

```{r YiYp_adj_model, echo = FALSE}
library(knitr)
knitr::include_graphics("assets/two_trait_model_yiyp_adj.png", error = FALSE)
```

# Results.

## Overlap between IVs for $X$ and $Y$.

- Is simply calculating the SNP overlap good enough? 
- Better to calculate signal overlap using LD at some $r^2$ cutoff (what threshold would be useful?).
  - For each IV used to estimate the causal effect $Y_i \rightarrow Y_p$,
  - Test if they are in LD with any IV used to estimate the causal effect $X_i \rightarrow X_p$ at $r^2$ > `0.9`, 
  - Repeat for all IVs for $Y_i$ to calcualte the percentage overlap between the two sets of IVs. 
  - Define percentage overlap as: (# of $Y_i$ IVs are in LD with at least one $X_i$ IV at $r^2$ > `0.9`) / (total number of $Y_i$ IVs used).

The results can be seen below. The `YiXi_IV_exact_overlap` column indicates the percentage of $Y_i$ IVs that are an exact match with $X_i$ IVs (same `rs#`). The `YiXi_IV_sig_overlap` column denotes the percentage overlap of signals, as defined above.

```{r XiYi_IV_overlap, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
proxyyMR_IV_overlap.DT <- proxyyMR_IV_overlap %>% dplyr::select(exposure_description, outcome_description, YiXi_IV_exact_overlap, YiXi_IV_sig_overlap) %>% unique() %>% arrange(desc(YiXi_IV_sig_overlap))

DT::datatable(proxyyMR_IV_overlap.DT, width = "100%") %>% formatRound(columns=c('YiXi_IV_exact_overlap', 'YiXi_IV_sig_overlap'), digits = 3)
```

The results with the most overlap involve traits that are nearly identical.

## Adjust $Y_i \rightarrow Y_p$.

Performed MV MR $Y_i \rightarrow Y_p$ relationships: $Y_p \sim Y_i + X_i$. Where IVs were the same as the two trait MR ($Y_p \sim Y_i$) based on $Y$, but the effect of the IVs on $X_i$ were also included in the model.

```{r YiYp_adj, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
yiyp_adj.DT <- proxyMR_MR_paths_summary_yiyp_adj %>% dplyr::select(exposure_description, outcome_description, exposure_sex, starts_with("yiyp_IVW"))

DT::datatable(yiyp_adj.DT, width = "100%") %>% formatRound(columns=c('YiXi_IV_exact_overlap', 'YiXi_IV_sig_overlap'), digits = 3)

fig_dat <- full_join(proxyyMR_IV_overlap, proxyMR_MR_paths_summary_yiyp_adj)

create_fig <- function(fig_dat, custom_col){

  #AM_result <- household_MR_binned_het %>% filter(same_trait)

  plot_data <- fig_dat %>% #dplyr::select(exposure_ID, exposure_sex, IVW_beta, IVW_se, IVW_pval, sex_het_pval) %>% unique() %>%
    #pivot_wider(names_from = "exposure_sex", values_from = c("IVW_beta", "IVW_se", "IVW_pval")) %>%
    mutate(overlap_above_50 = case_when(YiXi_IV_sig_overlap > 0.5 ~ TRUE, TRUE ~ FALSE))

  legend_title <- "IV overlap"
  plot <- ggplot2::ggplot(data = plot_data, ggplot2::aes(x = yiyp_IVW_beta_adj,
                                                         y = yiyp_IVW_beta, color = factor(overlap_above_50))) +
    geom_point(alpha = 3/4) +

    geom_smooth(mapping = aes(x = yiyp_IVW_beta_adj, y = yiyp_IVW_beta), method = "lm", se=FALSE, formula = y~x+0, fullrange=TRUE, color = custom_col[4]) +
    geom_abline(slope=1, intercept=0, color = "black") +
    scale_color_manual(values = custom_col[c(1,2)], labels=c("overlap > 50%","overlap =< 50%")) +
    theme_minimal() +
    ggplot2::labs(colour = legend_title, x = paste0("MR estimate (Y-index to Y-partner)"),
                  y = paste("AM MR estimate adjusted for IV effect on relevant X"))
}

```

## Updated results with adjusted $Y_i$

```{r proxy_MR_update, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

t1 <- as.data.frame(table(proxyMR_comparison_summary_yiyp_adj$omega_vs_gam_BF_sig_meta))
colnames(t1) <- c("omega_vs_gam_BF_sig_meta", "Frq")

kbl(t1) %>% kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

t2 <- as.data.frame(table(proxyMR_comparison_summary_yiyp_adj$omega_vs_rho_BF_sig_meta))
colnames(t2) <- c("omega_vs_rho_BF_sig_meta", "Frq")

kbl(t2) %>% kable_styling(bootstrap_options = "striped", full_width = F, position = "left")

t3 <- as.data.frame(table(proxyMR_comparison_summary_yiyp_adj$omega_vs_gam_rho_BF_sig_meta))
colnames(t3) <- c("omega_vs_gam_rho_BF_sig_meta", "Frq")

```

## MVMR

Run all MVMR models but need to sum up along all $\beta$ from the MV model to create the $\gamma_{mv}$.

MV MR: $X \rightarrow  Z \rightarrow Y$
  - Drop Z's if not significant in MV MR (i.e. $Y \sim X + Z_1 + Z_2 + Z_3...$)
  - Do we drop from the model itself and recompute the beta-coefficients? Or simply not include them in the result sum?

```{r MV_model, echo = FALSE}
library(knitr)
knitr::include_graphics("assets/MV_trait_model_yiyp_adj.png", error = FALSE)
``