### 2021-10-20 analysis/update_meeting_28_10_2021.Rmd
add update meeting for 28-10-21

---
title: "update_meeting_28_10_2021"
author: "jennysjaarda"
date: "2021-10-20"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}

options(scipen=999)
source("code/settings.R")
source("R/functions.R")
library(targets)
library(dplyr)
library(DT)
library(knitr)
library(ggplot2)
library(kableExtra)

```

```{r load_targets, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
tar_load(household_MR_summary)
tar_load(PCs_corr)
tar_load(PC_trait_corr)
tar_load(corr_impact_by_PCs)
tar_load(corr_impact_by_coords)
tar_load(proxyMR_MR_paths_summary_yiyp_adj)
tar_load(MV_z_corr_filter)
tar_load(traits_corr2_filled)
tar_load(proxyMR_prod_comparison_fig_yiyp_adj)
tar_load(proxyMR_prod_comparison_figure)

```

# Progress update.

## Compairing paths from $X_i \rightarrow Y_p$.

Without adjustment.

``````{r proxyMR_figure1, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600}
proxyMR_prod_comparison_figure
```

With adjustment.

``````{r proxyMR_figure2, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600}
proxyMR_prod_comparison_fig_yiyp_adj
```