### 2021-10-26 analysis/update_meeting_25_10_2021.Rmd
wflow_rename("analysis/update_meeting_28_10_2021.Rmd", "analysis/update_meeting_25_10_2021.Rmd")

---
title: "Update Meeting: 25-10-2021"
author: "jennysjaarda"
date: "2021-10-20"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}

options(scipen=999)
source("code/settings.R")
source("R/functions.R")
library(targets)
library(dplyr)
library(DT)
library(knitr)
library(ggplot2)
library(kableExtra)
library(cowplot)

```

```{r load_targets, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
tar_load(household_MR_summary)
tar_load(PCs_corr)
tar_load(PC_trait_corr)
tar_load(corr_impact_by_PCs)
tar_load(corr_impact_by_coords)
tar_load(proxyMR_MR_paths_summary_yiyp_adj)
tar_load(proxyMR_comparison_summary_yiyp_adj)
tar_load(MV_z_corr_filter)
tar_load(traits_corr2_filled)
tar_load(proxyMR_prod_comparison_fig_yiyp_adj)
tar_load(proxyMR_prod_comparison_fig)

```

# Progress update.

## Impact of geography.

Wanted to assess impact of geography on couple trait correlation. Investigated this by using both genetic PCs and birth coordinates.

### PC impact.

Tested the following:

- cor(PC_i,PC_p) for each PC (i.e. within couple PC correlation).
- cor(X,PC) for all PCs.

Next, I calculated the correlation due to counfounding as `cor(X,PC)^2*cor(PC_i,PC_p)` values, and plotted them against the raw `cor(X_i,X_p)` values.

```{r PC_trait_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

plot_data <- corr_impact_by_PCs %>% dplyr::select(Neale_pheno_ID, trait_couple_corr, corr_due_to_confounding_all) %>% unique()
plot <- ggplot2::ggplot(data = plot_data, ggplot2::aes(x = trait_couple_corr,
                                                       y = corr_due_to_confounding_all)) +
  geom_point(alpha = 3/4) +

  geom_smooth(mapping = aes(x = trait_couple_corr, y = corr_due_to_confounding_all), method = "lm", se=FALSE, formula = y~x+0, fullrange=TRUE, color = custom_col[4]) +
  theme_minimal() +
  ggplot2::labs(x = paste0("Couple trait correlation"),
                y = paste("Couple correlation due to confounding"))

plot

```

The table below gives the data in the plot above:

- `outcome_description` corresponds to the trait. 
- `trait_couple_corr` is the correlation of `outcome_description` in couples. 
-  `corr_due_to_confounding_all` corresponds to the formula: `cor(X,PC)^2*cor(PC_i,PC_p)`, summed across all PCs.

```{r PC_traits_confounding, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

corr_impact_by_PCs.DT <- replace_Neale_ID(plot_data, traits_corr2_filled, "Neale_pheno_ID", "outcome_description") %>% ungroup()
format_round_cols <- colnames(dplyr::select_if(corr_impact_by_PCs.DT, is.numeric))
DT::datatable(corr_impact_by_PCs.DT, width = "100%")  %>% formatRound(columns=format_round_cols, digits = 3)

```

### Coordiante impact.

Performed the same analysis as above but replacing PCs with North and East birth co-ordinates (data field [129](https://biobank.ctsu.ox.ac.uk/crystal/field.cgi?id=129) and [130](https://biobank.ctsu.ox.ac.uk/crystal/field.cgi?id=130), respectively).

```{r coord_trait_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

plot_data <- corr_impact_by_coords %>% dplyr::select(Neale_pheno_ID, trait_couple_corr, corr_due_to_confounding_all) %>% unique()
plot <- ggplot2::ggplot(data = plot_data, ggplot2::aes(x = trait_couple_corr,
                                                       y = corr_due_to_confounding_all)) +
  geom_point(alpha = 3/4) +

  geom_smooth(mapping = aes(x = trait_couple_corr, y = corr_due_to_confounding_all), method = "lm", se=FALSE, formula = y~x+0, fullrange=TRUE, color = custom_col[4]) +
  theme_minimal() +
  ggplot2::labs(x = paste0("Couple trait correlation"),
                y = paste("Couple correlation due to confounding"))

plot

```

The table below gives the data in the plot above:

- `outcome_description` corresponds to the trait. 
- `trait_couple_corr` is the correlation of `outcome_description` in couples. 
-  `corr_due_to_confounding_all` corresponds to the formula: `cor(X,PC)^2*cor(PC_i,PC_p)`, summed across both coorindates.

```{r coord_traits_confounding, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

corr_impact_by_coords.DT <- replace_Neale_ID(plot_data, traits_corr2_filled, "Neale_pheno_ID", "outcome_description") %>% ungroup()
format_round_cols <- colnames(dplyr::select_if(corr_impact_by_coords.DT, is.numeric))
DT::datatable(corr_impact_by_coords.DT, width = "100%")  %>% formatRound(columns=format_round_cols, digits = 3)

```

## Compairing paths from $X_i \rightarrow Y_p$.

We performed two analyses to compare $\rho$, $\gamma$ and $\omega$:

- No adjustment in $Y_i \rightarrow Y_p$ MR. 
- Adjustment for effects on $X_i$ in $Y_i \rightarrow Y_p$ MR,  MVMR model as follows: $Y_p \sim Y_i + X_i$ (IVs were a combination of all $X$ and $Y$ IVs, pruned at standard parameters using 1000G European data).

```{r YiYp_adj_model, echo = FALSE}
library(knitr)
knitr::include_graphics("assets/two_trait_model_yiyp_adj.png", error = FALSE)
```

Broad overview of results are shown in the figure below.

- Without adjustment.

```{r proxyMR_figure_no_adj, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600}
proxyMR_prod_comparison_fig
```

- With adjustment.

```{r proxyMR_figure_adj, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600}
proxyMR_prod_comparison_fig_yiyp_adj
```

Going forward, we will just use the adjusted results.

### Exploring the difference between $\rho$ and $\gamma$.

In general, $\rho$ is significantly larger that $\gamma$, meaning that $X_i \rightarrow Y_p$ favors paths where assortative mating is through $X$ rather than $Y$. This makes a lot of biological and intuitive sense. In other words, exposures are passed from index to partner, rather than outcomes.

A summary of the linear model of $\rho$ vs $\gamma$, forced through the intercept, is below.

```{r proxyMR_summary_rho_vs_gam, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

data <- proxyMR_comparison_summary_yiyp_adj
exposure_sex <- "meta"

fig_data <- data[which(data[["exposure_sex"]]=="male"),]

x_start <- "gam"
y_start <- "rho"
overlay_var_start <- "gam_vs_rho_BF_sig"
overlay_var = paste0(overlay_var_start, "_meta")

x = paste0(x_start, "_beta")
y = paste0(y_start, "_beta")

count <- 0

fig_data <- fig_data %>% rename(x := !!x) %>% rename(y := !!y) %>% rename(overlay_var := !!overlay_var) %>%
    mutate(x_plot = case_when(x < 0 ~ abs(x), TRUE ~ x)) %>%
    mutate(y_plot = case_when(x < 0 ~ -1*y, TRUE ~ y))

summary(lm(y ~ x + 0, data = fig_data))

```

The corresponding figure is below. The blue line includes only BF-significant $\rho$ and $\gamma$, where the green line includes all points (analagous to the linear model above).

```{r proxyMR_fig_rho_vs_gam, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600, message = FALSE, warning = FALSE}

plot <- create_proxy_prod_comparison_fig_ind(data, exposure_sex, x, y, overlay_var, count)

plot[[1]]

```

### Exploring the difference between $\omega$ and [$\rho$ and $\gamma$].

A few observations:

- In general, $\omega$ is significantly larger than both $\rho$ and $\gamma$, as shown in the figures below. 
- On the other hand, when we sum $\rho$ and $\gamma$, $\omega$ is in general lower than the sum. 
- There are a number of instances where $\omega$ is significantly **smaller** than the sum &rightarrow not as interesting? *But what does this mean?* That we didn't have power in the $\omega$ (i.e diagonal) MR to capture the relationship that is likely only going through the either $\rho$ or $\gamma$ (or a sum of the two)? 
- There are a few cases where $\omega$ is significantly **larger** than the sum. See below for details.

```{r proxyMR_fig_omega_vs_gam, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600, message = FALSE, warning = FALSE}
data <- proxyMR_comparison_summary_yiyp_adj
exposure_sex <- "meta"

fig_data <- data[which(data[["exposure_sex"]]=="male"),]

x_start <- "gam"
y_start <- "omega"
overlay_var_start <- "omega_vs_gam_BF_sig"
overlay_var = paste0(overlay_var_start, "_meta")

x = paste0(x_start, "_beta")
y = paste0(y_start, "_beta")

count <- 0

plot <- create_proxy_prod_comparison_fig_ind(data, exposure_sex, x, y, overlay_var, count)

plot[[1]]

```

```{r proxyMR_fig_omega_vs_rho, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600, message = FALSE, warning = FALSE}
data <- proxyMR_comparison_summary_yiyp_adj
exposure_sex <- "meta"

fig_data <- data[which(data[["exposure_sex"]]=="male"),]

x_start <- "rho"
y_start <- "omega"
overlay_var_start <- "omega_vs_rho_BF_sig"
overlay_var = paste0(overlay_var_start, "_meta")

x = paste0(x_start, "_beta")
y = paste0(y_start, "_beta")

count <- 0

plot <- create_proxy_prod_comparison_fig_ind(data, exposure_sex, x, y, overlay_var, count)

plot[[1]]

```

```{r proxyMR_fig_omega_vs_gam_rho, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600, message = FALSE, warning = FALSE}
data <- proxyMR_comparison_summary_yiyp_adj
exposure_sex <- "meta"

fig_data <- data[which(data[["exposure_sex"]]=="male"),]

x_start <- "gam_rho"
y_start <- "omega"
overlay_var_start <- "omega_vs_gam_rho_BF_sig"
overlay_var = paste0(overlay_var_start, "_meta")

x = paste0(x_start, "_beta")
y = paste0(y_start, "_beta")

count <- 0

plot <- create_proxy_prod_comparison_fig_ind(data, exposure_sex, x, y, overlay_var, count)

plot[[1]]

 fig_data <- fig_data %>% rename(x := !!x) %>% rename(y := !!y) %>% rename(overlay_var := !!overlay_var) %>%
    mutate(x_plot = case_when(x < 0 ~ abs(x), TRUE ~ x)) %>%
    mutate(y_plot = case_when(x < 0 ~ -1*y, TRUE ~ y))

 summary(lm(y ~ x + 0, data = fig_data))

```

A summary of the cases where $\omega$ is significantly larger or smaller than the sum of $\rho$ and $\gamma$ are shown in the two tables below (taking absolute values of each).

```{r proxyMR_summary_omega_vs_gam_rho, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600, message = FALSE, warning = FALSE}

omega_larger.DT <- proxyMR_comparison_summary_yiyp_adj %>% filter(omega_vs_gam_rho_BF_sig_meta) %>% filter(abs(omega_meta_beta) > abs(gam_rho_meta_beta)) %>% dplyr::select(exposure_description, outcome_description, gam_rho_meta_beta, omega_meta_beta) %>% unique() 
format_round_cols <- colnames(dplyr::select_if(omega_larger.DT, is.numeric))
DT::datatable(omega_larger.DT, width = "100%")  %>% formatRound(columns=format_round_cols, digits = 3)

omega_smaller.DT <- proxyMR_comparison_summary_yiyp_adj %>% filter(omega_vs_gam_rho_BF_sig_meta) %>% filter(abs(gam_rho_meta_beta) > abs(omega_meta_beta)) %>% dplyr::select(exposure_description, outcome_description, gam_rho_meta_beta, omega_meta_beta) %>% unique() 
format_round_cols <- colnames(dplyr::select_if(omega_smaller.DT, is.numeric))
DT::datatable(omega_smaller.DT, width = "100%")  %>% formatRound(columns=format_round_cols, digits = 3)

#fig_data %>% filter(omega_meta_beta < 0 & gam_rho_meta_beta > 0) %>% dplyr::select(x,y, exposure_description, outcome_description, overlay_var)

```

## Next steps.

- In the $Y_i \rightarrow Y_p$ MR, rather than performing a MVMR ($Y_p \sim Y_i + X_i$), try in a two step process as follows: 
    - Calculate the effect of $Y_p$ without effects of $X_i$ on $Y_p$ as: $Y_{resid} = Y_p - \omega*X_I$.
    - Run MR with the residualized $Y$ as: $Y_{resid} \sim Y_i$

**Question**: What is the corresponding SEs on $Y_{resid}$?

- Check for cases of reverse causation in the same-person MR. 
- Remove SNPs were the effect is significantly larger on the outcome than the exposure. 
- In the results including both sexes, meta-analyze at the SNP-level rather than the MR level (note that right now the two horizontal, same person MRs in the DAG above are different because they are sex specific. I then meta-analyze $\omega$, $\gamma$ and $\rho$ as the *final* step).
- Calculate the variance of $\rho + \gamma$ as: $2 * cor(\rho, \gamma) * SE_{\gamma} * SE{\rho}$, i.e. include the extra-term of 2*correlation.

# PolyMR.

- Finished discussion points. 
- Waiting on Jonathan to add bibtex file before adding my references. 
- Haven't done a summary on CATE - who was going to do that? 
- *Question:* Why is there often an inflection point around the (0,0) point? Is this an effect of standardization?