### 2021-11-03 analysis/update_meeting_03_11_2021.Rmd
wflow_publish("analysis/update_meeting_03_11_2021.Rmd")

---
title: "Update Meeting: 03-11-2021"
author: "Jenny Sjaarda"
date: "2021-11-01"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}

options(scipen=999)
source("code/settings.R")
source("R/functions.R")
library(targets)
library(dplyr)
library(DT)
library(knitr)
library(ggplot2)
library(kableExtra)

```

```{r load_targets, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
tar_load(household_MVMR_summary)
tar_load(household_MR_summary)

```

# Progress update

Decided to run MVMR as follows: $Y_p \sim X_i + Y_i + X_p$, using instruments for ${X_i, Y_i, X_p}$. In this MR, the coefficient for $X_i$ would represent the *direct* $X_i \rightarrow Y_p$ causal effect and hopefully this would be close to zero in most cases. Running simply the $Y_p \sim X_i$ MR (with only $X_i$ instruments), would give you the $X_i \rightarrow Y_p$ *total* effect. Then we can compare the direct and total effect results.

A summary of the MVMR results are below, filtered to only traits with abs(correlation) < 0.8, and then BF significant with $X_i$.

The column `corr_traits` corresponds to the raw correlation between traits in the biobank. The columns `IVW_meta_beta` and `IVW_meta_pval` correspond to the meta-analyzed across sexes univariate results from the $Y_p ~ X_i$ MR. The subsequent columns correspond to the MVMR MR results.

```{r MVMR_summary, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}

household_MVMR_summary.DT <- household_MVMR_summary %>% filter(same_trait==FALSE) %>% filter(abs(corr_traits) < 0.8)

num_tests <- dim(household_MVMR_summary.DT)[1]
household_MVMR_BF.DT <- household_MVMR_summary.DT %>% filter(xi_meta_pval < 0.05/num_tests) %>% dplyr::select(exposure_ID, outcome_ID, exposure_description, outcome_description, corr_traits, yi_meta_beta, yi_meta_pval, xi_meta_beta, xi_meta_pval, xp_meta_beta, xp_meta_pval) %>% unique()

household_MR_summary_sub <- household_MR_summary %>% dplyr::select(exposure_ID, outcome_ID, IVW_meta_beta, IVW_meta_se, IVW_meta_pval) %>% unique()
MVMR_uniMR_comparison.DT <- inner_join(household_MR_summary_sub, household_MVMR_BF.DT)  %>% dplyr::select(exposure_ID, outcome_ID, exposure_description, outcome_description, corr_traits, everything())
format_round_cols <- colnames(dplyr::select_if(MVMR_uniMR_comparison.DT, is.numeric))

DT::datatable(MVMR_uniMR_comparison.DT, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3)

```

```{r direct_vs_indirect_effects_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}

plot_data <-  inner_join(household_MR_summary_sub, household_MVMR_summary.DT)  %>% dplyr::select(exposure_ID, outcome_ID, exposure_description, outcome_description, corr_traits, IVW_meta_beta, IVW_meta_se, yi_meta_beta, yi_meta_se, xi_meta_beta, xi_meta_se, xp_meta_beta, xp_meta_se) %>% unique()

#bquote('x axis'~(Å^2))
#expression(Sepal~Length[cm])

#Indirect~effect~of~X[i]~on~Y[p]~(from~univariate~MR)
#xlab <- bquote('Indirect effect of x_i on y_p (from univariate MR)'~(Å^2))

plot <-
  ggplot2::ggplot(data = plot_data, ggplot2::aes(x = IVW_meta_beta,
                                                 y = xi_meta_beta)) +
  geom_point(alpha = 3 / 4) +
  geom_smooth(
    mapping = aes(x = IVW_meta_beta, y = xi_meta_beta),
    method = "lm",
    se = FALSE,
    formula = y ~ x + 0,
    fullrange = TRUE,
    color = custom_col[4]
  ) +
  ggplot2::geom_errorbar(
    ggplot2::aes(ymin = xi_meta_beta - xi_meta_se, ymax = xi_meta_beta + xi_meta_se),
    colour = "grey",
    width = 0
  ) +
  ggplot2::geom_errorbarh(
    ggplot2::aes(xmin = IVW_meta_beta - IVW_meta_se, xmax = IVW_meta_beta + IVW_meta_se),
    colour = "grey",
    height = 0
  ) +
  theme_minimal() +
  ggplot2::labs(
    x = expression(Indirect~effect~of~X[i]~on~Y[p]~(from~univariate~MR)),
    y = expression(Direct~effect~of~X[i]~on~Y[p]~(from~MVMR))
  ) + 
  geom_abline(slope=1, intercept=0)

plot

```