### 2022-01-12 analysis/manuscript.Rmd
create a manuscript Rmd with figures

---
title: "Manuscript"
author: "Jenny Sjaarda"
date: "2022-01-04"
output: workflowr::wflow_html
---

```{r setup, include = FALSE}

options(scipen=999)
source("code/settings.R")
source("R/functions.R")
library(targets)
library(kableExtra)
library(dplyr)
library(DT)
library(knitr)
library(ggplot2)
library(kableExtra)
library(cowplot)
library(plotly)
library(plyr)
```

```{r load_targets, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
tar_load(couple_MR_vs_trait_corr)
tar_load(corr_impact_by_traits)
tar_load(traits_corr2_filled)
tar_load(household_MR_summary_AM)
tar_load(num_tests_by_PCs)
tar_load(num_tests_by_PCs_AM_sig)
tar_load(household_MR_binned_het_sex_spec)
tar_load(household_MR_binned_het_joint)
tar_load(traits_final)
tar_load(household_MR_AM_FvsM_fig)
tar_load(confounder_traits)

```

# A few notes.

- The number of effective tests can be found at the target `num_tests_by_PCs`. 
- The number of effective tests among those significant in the MR can be found at the target `num_tests_by_PCs_AM_sig`.

# Relationship between causal effects and raw phenotypic correlation in couples

Sought to compare the raw correlation amongst couples vs. the standardized MR effects. There are many traits where the `correlation > MR effects` including some traits of interest such as standing height, place of birth, among many others. A plot comparing the two and a summary table of the significant effects is shown below.

```{r corr_vs_MR_effects, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}

plot_data <- replace_Neale_ID(couple_MR_vs_trait_corr, traits_corr2_filled, "exposure_ID", "outcome_description") %>% 
  dplyr::rename(Phenotype = outcome_description) %>% 
  dplyr::rename(Correlation = couple_r) %>% dplyr::rename(MR_est = IVW_beta)

plot <- ggplot2::ggplot(data = plot_data, ggplot2::aes(x = Correlation,
                                                       y = MR_est, Phenotype = Phenotype)) +
  geom_point(alpha = 3/4) +

  geom_abline(slope=1, intercept=0) +

  ggplot2::geom_errorbar(
    ggplot2::aes(ymin = MR_est - IVW_se, ymax = MR_est + IVW_se),
    colour = "grey",
    width = 0
  ) +
  ggplot2::geom_errorbar(
    ggplot2::aes(xmin = Correlation - couple_r_se, xmax = Correlation + couple_r_se),
    colour = "grey",
    width = 0
  ) +

  theme_minimal() +
  ggplot2::labs(x = paste0("Couple correlation"),
                y = paste("Standardized MR effect"))

ggplotly(plot, tooltip = c("x", "y", "Phenotype"))

```

```{r corr_vs_MR_effects_table, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}

plot_data.DT <- plot_data %>% dplyr::rename(IVW_beta = MR_est) %>% 
  filter(diff_p < 0.05/num_tests_by_PCs) %>%
  mutate(mr_corr_diff = Correlation - IVW_beta) %>% 
  mutate(correlation_larger = ifelse(Correlation > IVW_beta, TRUE, FALSE)) %>%
  dplyr::select(Phenotype, Correlation, IVW_beta, IVW_pval, mr_corr_diff, correlation_larger)

format_round_cols <- colnames(dplyr::select_if(plot_data.DT, is.numeric))

DT::datatable(plot_data.DT, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3)
```

# Effect of sex, age and time together on causal effects in couples

```{r AM_data_prep, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE, include = FALSE}

AM_result_sig <- household_MR_summary_AM %>% filter(IVW_pval < 0.05/num_tests_by_PCs) 
AM_sig_traits <-household_MR_summary_AM %>% filter(IVW_pval < 0.05/num_tests_by_PCs) %>% pull(exposure_ID) %>% unique()
AM_sex_spec_result_sig <- household_MR_binned_het_sex_spec %>% filter(same_trait ==TRUE) %>% filter(exposure_ID %in% AM_sig_traits) %>% filter(sex_het_pval < 0.05/num_tests_by_PCs_AM_sig)

```

- The number of analyses corresponds to the number of traits (`r dim(traits_final)[1]`). 
- The significant results after adjusting for multiple hypothesis testing are shown below. 
- `IVW_meta_beta` and `IVW_meta_pval` corresponds to the beta and p-value of the meta-analyzed MR across sexes, respectively (i.e. MR estimates were computed in each sex-seperately using sex-specific SNP-exposure and SNP-outcome results and then meta-analyzed).  
- After adjusting for multiple tests (p < `0.05/`r num_tests_by_PCs``), identified `r dim(AM_result_sig)[1]` significant assortative mating MR results (corresponding to the table below).

# Impact of confounders on couple correlation

We tested the impact of confounding on genetic PCs, N/E coordinates and the following 4 traits:

```{r confounder_traits, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
confounder_trait_tibble <- as_tibble(confounder_traits)
confounder_trait_tibble_descript <- replace_Neale_ID(confounder_trait_tibble, traits_corr2_filled, "value", "outcome_description")
confounder_trait_tibble_descript %>% 
  kbl() %>%
  kable_styling()
```

```{r confounder_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}

plots <- list()
for(i in 1:length(confounder_traits)){

  trait <- confounder_traits[i]
  trait_description <- confounder_trait_tibble_descript[[1]][i]

  plot_data <- corr_impact_by_traits %>% filter(trait_interest==trait) %>% filter(trait_interest!=trait_j)
  plot_data <- replace_Neale_ID(plot_data, traits_corr2_filled, "trait_j", "Phenotype")
  plot <-
    ggplot2::ggplot(
      data = plot_data,
      ggplot2::aes(x = trait_j_couple_corr,
                   y = corr_due_to_confounding, Phenotype = Phenotype)
    ) +
    geom_point(alpha = 3 / 4) +

    ggplot2::geom_errorbar(
      ggplot2::aes(ymin = corr_due_to_confounding - corr_due_to_confounding_se, ymax = corr_due_to_confounding + corr_due_to_confounding_se),
      colour = "grey",
      width = 0
    ) +
    ggplot2::geom_errorbar(
      ggplot2::aes(xmin = trait_j_couple_corr - trait_j_couple_corr_se, xmax = trait_j_couple_corr + trait_j_couple_corr_se),
      colour = "grey",
      width = 0
    ) +

    theme_minimal() +
    ggplot2::labs(
      x = paste0("Couple trait correlation"),
      y = paste0("Couple correlation due to confounding"),
      title = paste0("Impact of ", trait_description, "\n on couple correlation")
    )

  plots[[i]] <- plot 
}

figures_grid <- plot_grid(plotlist = plots, nrow=2, ncol = 2, byrow = F, labels="AUTO")
figures_grid
```