### 2022-02-03 analysis/manuscript.Rmd
wflow_publish("analysis/manuscript.Rmd")

---
title: "Manuscript"
author: "Jenny Sjaarda"
date: "2022-01-04"
output: workflowr::wflow_html
---

```{r setup, include = FALSE}

options(scipen=999)
source("code/settings.R")
source("R/functions.R")
library(targets)
library(kableExtra)
library(DT)
library(knitr)
library(ggplot2)
library(kableExtra)
library(cowplot)
library(plotly)
library(plyr)
library(dplyr)
library(ggrepel)
library(viridis)
library(hrbrthemes)
library(ggpubr)
library(broom)
```

```{r load_targets, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}
tar_load(couple_MR_vs_trait_corr)

tar_load(corr_impact_by_PCs)
tar_load(corr_impact_by_traits)
tar_load(corr_impact_by_coords)
tar_load(corr_potential_trait_confounders)
tar_load(traits_corr)
tar_load(PCs_corr)
tar_load(traits_corr2_filled)
tar_load(household_MR_summary_AM)
tar_load(num_tests_by_PCs)
tar_load(num_tests_by_PCs_AM_sig)
tar_load(household_MR_binned_het_sex_spec)
tar_load(household_MR_binned_het_SNPmeta)
tar_load(traits_final)
tar_load(household_MR_AM_FvsM_fig)
tar_load(confounder_traits)
tar_load(proxyMR_prod_comparison_fig_yiyp_adj)
tar_load(proxyMR_comparison_summary_yiyp_adj_SNPmeta)
tar_load(household_MR_summary_AM)
tar_load(household_MR_binned_het_SNPmeta)

```

# A few notes (for myself).

- The number of effective tests can be found at the target `num_tests_by_PCs`. 
- The number of effective tests among those significant in the MR can be found at the target `num_tests_by_PCs_AM_sig`.

# Relationship between causal effects and raw phenotypic correlation in couples.

Sought to compare the raw correlation amongst couples vs. the standardized MR effects. There are many traits where the `correlation > MR effects` including some traits of interest such as standing height, place of birth, among many others. A plot comparing the two and a summary table of the significant effects is shown below.

```{r corr_vs_MR_effects, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}

interesting_phenotypes <- c("Basal metabolic rate", "Place of birth in UK - north co-ordinate", 
                            "Systolic blood pressure, automated reading", "Standing height", 
                            "Body fat percentage", "Forced vital capacity (FVC), Best measure","Red blood cell (erythrocyte) count")

plot_data <- replace_Neale_ID(couple_MR_vs_trait_corr, traits_corr2_filled, "exposure_ID", "outcome_description") %>% 
  mutate(Phenotype = outcome_description) %>% 
  mutate(Phenotype = case_when(Phenotype == "Basal metabolic rate" ~ "BMR",
                               Phenotype == "Systolic blood pressure, automated reading" ~ "SBP",
                               Phenotype == "Body fat percentage" ~ "BFP",
                               Phenotype == "Forced vital capacity (FVC), Best measure" ~ "FVC",
                               Phenotype == "Red blood cell (erythrocyte) count" ~ "RBC count",
                               Phenotype == "Place of birth in UK - north co-ordinate" ~ "Place of birth, NC",
                               TRUE ~ Phenotype)) %>%
  dplyr::rename(Correlation = couple_r) %>% dplyr::rename(MR_est = IVW_beta) %>% 
  mutate(diff_p_sig = case_when(diff_p < 0.05/num_tests_by_PCs ~ TRUE, TRUE ~ FALSE)) %>%
  mutate(label_phenos = case_when(outcome_description %in% interesting_phenotypes ~ TRUE, TRUE ~ FALSE))

  # TRUE - BF sig difference 
  # FALSE - non-sig 
  #mutate(diff_p_sig = ifelse(diff_p < 0.05/num_tests_by_PCs, TRUE, FALSE))

plot <- ggplot2::ggplot(data = plot_data, ggplot2::aes(x = Correlation,
                                                       y = MR_est, 
                                                       label = Phenotype, 
                                                       color = factor(diff_p_sig)),) +
  geom_point(alpha = 1.2) +
  scale_color_manual(values = custom_col[c(1,2)], labels=c("Non-significant","BF-signifncant difference")) +
  #geom_text(show.legend = FALSE) +
  geom_abline(slope=1, intercept=0) +

  ggplot2::geom_errorbar(
    ggplot2::aes(ymin = MR_est - IVW_se, ymax = MR_est + IVW_se),
    colour = "grey",
    width = 0
  ) +
  ggplot2::geom_errorbar(
    ggplot2::aes(xmin = Correlation - couple_r_se, xmax = Correlation + couple_r_se),
    colour = "grey",
    width = 0
  ) +

  theme_minimal() +
  ggplot2::labs(x = paste0("Couple correlation"),
                y = paste("Standardized MR effect")) +
  theme(legend.position="bottom") + theme(legend.title=element_blank()) +
  geom_text_repel(data = subset(plot_data, label_phenos==TRUE),
                  nudge_y = -0.25 - subset(plot_data, label_phenos==TRUE)$MR_est, size = 2,
                  direction = "x")

plot

plot_data.DT <- plot_data %>% dplyr::rename(IVW_beta = MR_est) %>% 
  filter(diff_p < 0.05/num_tests_by_PCs) %>%
  mutate(mr_corr_diff = Correlation - IVW_beta) %>% 
  mutate(correlation_larger = ifelse(Correlation > IVW_beta, TRUE, FALSE)) %>%
  dplyr::select(outcome_description, Correlation, IVW_beta, IVW_pval, diff_p, mr_corr_diff)

```

## Summary of cases where correlation > MR estimate.

The table below displays the `r dim(plot_data.DT)[1]` traits that have correlation significantly > MR-estimate.

```{r corr_vs_MR_effects_table, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}

format_round_cols <- colnames(dplyr::select_if(plot_data.DT, is.numeric))

DT::datatable(plot_data.DT, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3)
```

## Search for potential confounders to explain cases where correlation > MR estimate.

1. Restrict to only traits that have correlation significantly > MR-estimate (`r dim(plot_data.DT)[1]` traits have correlation > MR_estimate).
2. For each of these traits, find the traits which have a causal effect on this trait (MR p-val < 0.05/[number of effective tests]).
3. Calculate the correlation due to confounding as: $C = (\alpha_{z\rightarrow x})^2 * \alpha_{z_i\rightarrow z_p}$.
4. Calculate the ratio of this correlation ($C$) and the correlation of $X$ in partners ($C/r_x$).

A summary of the these results can be seen below:

```{r confounder_summary, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}

number_of_potential_confounders <- dim(corr_potential_trait_confounders)[1]/dim(plot_data.DT)[1]

non_sig_confounders <- corr_potential_trait_confounders %>% filter(!sig_confounder) %>% group_by(outcome_ID) %>%
  dplyr::summarize(description = outcome_description[1], n = n()) %>% 
  filter(n==number_of_potential_confounders) %>%
  mutate(n = n-number_of_potential_confounders) %>% mutate(potential_confounders = NA) %>%
  mutate(mean_C_ratio = NA) %>%
  dplyr::rename(trait_ID = outcome_ID)

sig_confounders <- corr_potential_trait_confounders %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
  #mutate(potential_confounders = paste0(exposure_description, collapse = "; ")) %>%
  dplyr::summarize(description = outcome_description[1], n = n(), 
                   potential_confounders = paste0(exposure_description, collapse = "; "), 
                   mean_C_ratio = mean(corr_due_to_confounding_ratio)) %>%
  dplyr::rename(trait_ID = outcome_ID)

confounder_summary <- rbind(non_sig_confounders, sig_confounders)

format_round_cols <- colnames(dplyr::select_if(confounder_summary, is.numeric))

DT::datatable(confounder_summary, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3)

```

# Effect of sex, age and time together on causal effects in couples.

```{r AM_data_prep, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE, include = FALSE}

AM_result_sig <- household_MR_summary_AM %>% filter(IVW_pval < 0.05/num_tests_by_PCs) 
AM_sig_traits <-household_MR_summary_AM %>% filter(IVW_pval < 0.05/num_tests_by_PCs) %>% pull(exposure_ID) %>% unique()
AM_sex_spec_result_sig <- household_MR_binned_het_sex_spec %>% filter(same_trait ==TRUE) %>% filter(exposure_ID %in% AM_sig_traits) %>% filter(sex_het_pval < 0.05/num_tests_by_PCs_AM_sig)

```

- For each trait in the pipeline, we tested the causal effect within couples (i.e. the number of analyses corresponds to the number of traits (`r dim(traits_final)[1]`)). 
- The significant results after adjusting for multiple hypothesis testing are shown below. 
- `IVW_meta_beta` and `IVW_meta_pval` corresponds to the beta and p-value of the meta-analyzed MR across sexes, respectively (i.e. MR estimates were computed in each sex-seperately using sex-specific SNP-exposure and SNP-outcome results and then meta-analyzed).  
- After adjusting for multiple tests (p < `0.05/`r num_tests_by_PCs``), identified `r dim(AM_result_sig)[1]` significant assortative mating (same-trait) MR results (corresponding to the table below)..

```{r, AM_sig, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE, include = TRUE}

AM_result_sig.DT <- AM_result_sig %>% dplyr::select(exposure_description, IVW_beta, IVW_pval) %>% unique()
DT::datatable(AM_result_sig.DT, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval'), digits=3)

```

## Sex heterogeneity.

- For each AM MR ($X_{p} \sim X_i$), analyses were run in each sex separately. 
- Of the `r dim(AM_sig_traits)[1]` significant results shown above, we tested to see if there was any difference in AM MR estimates between sexes. 
- A similar adjustment for multiple hypothesis testing was applied as above, resulting in `r num_tests_by_PCs_AM_sig` number of effective tests. 
- After adjusting for multiple hypothesis testing (p < `0.05/`r num_tests_by_PCs_AM_sig``), `r dim(AM_sex_spec_result_sig)[1]` traits showed significant differences amongst sexes (this changed from before).
- The table below shows the *nominally significant* results at `p < 0.05`.

```{r AM_sex_het_nom, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}

AM_sig_sex_het_nom <- household_MR_binned_het_sex_spec %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% filter(sex_het_pval < 0.05)
AM_sig_sex_het_nom <- replace_Neale_ID(AM_sig_sex_het_nom, traits_corr2_filled,"exposure_ID", "exposure_description")
AM_result_sig_nom.DT <- AM_sig_sex_het_nom %>% dplyr::select(exposure_description, exposure_sex, IVW_beta, IVW_pval, sex_het_pval) %>% unique()
DT::datatable(AM_result_sig_nom.DT, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval', 'sex_het_pval'), digits=3)

```

## Binned results (exploring effect of age and time-together).

```{r binned_results_prep, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE, include = FALSE}

slope_sig_age <- household_MR_binned_het_SNPmeta %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
  filter(grouping_var == "age_even_bins") %>% 
  filter(bin_slope_pval < 0.05/num_tests_by_PCs_AM_sig)

slope_sig_tt <- household_MR_binned_het_SNPmeta %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
  filter(grouping_var == "time_together_even_bins") %>% 
  filter(bin_slope_pval < 0.05/num_tests_by_PCs_AM_sig)

slope_wt_sig_age <- household_MR_binned_het_SNPmeta %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
  filter(grouping_var == "age_even_bins") %>% 
  filter(bin_slope_pval_wt < 0.05/num_tests_by_PCs_AM_sig)

slope_wt_sig_tt <- household_MR_binned_het_SNPmeta %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
  filter(grouping_var == "time_together_even_bins") %>% 
  filter(bin_slope_pval_wt < 0.05/num_tests_by_PCs_AM_sig)

sex_het_sig_age <- household_MR_binned_het_sex_spec %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
  filter(grouping_var == "age_even_bins") %>% 
  filter(bin_slope_sex_het_pval < 0.05/num_tests_by_PCs_AM_sig)

sex_het_sig_tt <- household_MR_binned_het_sex_spec %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
  filter(grouping_var == "time_together_even_bins") %>% 
  filter(bin_slope_sex_het_pval < 0.05/num_tests_by_PCs_AM_sig)

sex_het_wt_sig_age <- household_MR_binned_het_sex_spec %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
  filter(grouping_var == "age_even_bins") %>% 
  filter(bin_slope_sex_het_pval_wt < 0.05/num_tests_by_PCs_AM_sig)

sex_het_wt_sig_tt <- household_MR_binned_het_sex_spec %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
  filter(grouping_var == "time_together_even_bins") %>% 
  filter(bin_slope_sex_het_pval_wt < 0.05/num_tests_by_PCs_AM_sig)

sex_het_wt_sig_age_traits <- replace_Neale_ID(sex_het_wt_sig_age, traits_corr2_filled, "exposure_ID", "outcome_description") %>% pull(outcome_description) %>% unique()
```

- For each AM MR ($X_{p} \sim X_i$), analyses were run in the full sample as well as in 5 roughly equal sized bins according to time couple had been together (`time_together_even_bins`) estimated using [time at household variable](https://biobank.ctsu.ox.ac.uk/showcase/field.cgi?id=699), and median age (`age_even_bins`). 
- Of the `r length(AM_sig_traits)` significant results shown above, we tested to see if there was any significant difference in AM MR estimates amongst the two grouping variables. 
- Specifically, the outcome data set was split into 5 bins, and the SNP-outcome effect was estimated in each bin separately (and each sex seperately). 
- These outcome-SNP effects were then used to generate bin-specific MR estimates, using the same SNP-exposure effects from Neale. 
- To estimate the difference among bins, we used two approaches: 
    1. The Cochran's Q test to test for heterogeneity in meta-analyses *(of which we obtained no significant results, so not presenting in paper)*.
    2. Tested the significance of the slope of linear model of median bin (either `age` or `time together`) versus the bin-specific MR estimate.

------------------------------------------------------------------------

Slopes were calculated by estimating the beta-coefficient of a linear regression between MR estimate within each bin (dependent variable) and median bin (independent variable, either age or time at same household). Linear models were run both *unweighted* and *weighted* for the by the inverse of the SE of the MR estimate.

The number of significant trends ($\beta$ estimates from the model: $\alpha_{bin} \sim median_{bin}$) significant after multiple hypothesis testing (p < `0.05/`r num_tests_by_PCs_AM_sig` = `r 0.05/num_tests_by_PCs_AM_sig``) in each group was:

  - Binned by median *age* of couples, number of significant results after adjusting for number of tests:
      - $\beta$ unweighted, meta-anlayzed across sexes: `r dim(slope_sig_age)[1]`
      - $\beta$ weighted by inverse of the SE, meta-anlayzed across sexes: `r dim(slope_wt_sig_age)[1]`
      - $\beta$ unweighted, significantly different between sexes: `r dim(sex_het_sig_age)[1]/2`
      - $\beta$ weighted by inverse of the SE, significantly different between sexes: `r dim(sex_het_wt_sig_age)[1]/2`

  - Binned by *time* couple had been together, number of significant results after adjusting for number of tests:
      - $\beta$ unweighted, meta-anlayzed across sexes: `r dim(slope_sig_tt)[1]`
      - $\beta$ weighted by inverse of the SE, meta-anlayzed across sexes: `r dim(slope_wt_sig_tt)[1]`
      - $\beta$ unweighted, significantly different between sexes: `r dim(sex_het_sig_tt)[1]/2`.
      - $\beta$ weighted by inverse of the SE, significantly different between sexes: `r dim(sex_het_wt_sig_tt)[1]/2`

Given that we have no significant results, how should we report these analyses?

A few potential figures we could include for a given trait are shown below. 
```{r binned_figures, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE, include = TRUE}

tar_load(exposure_info)
for(i in 1:length(exposure_info)){
  temp <- exposure_info[[i]]
  if(temp[which(temp$Value=="trait_ID"), "Info"] == "130_irnt") {break}
}

figs <- tar_read(household_MR_binned_AM_figs, branches = i)[[1]]

figs[[1]]
figs[[2]]
figs[[3]]
figs[[4]]
figs[[5]]
figs[[6]]
figs[[7]]
figs[[8]]
figs[[9]]

```

# Heterogeneity stats

- In the single trait analysis, we also decided to investigate Cochran’s heterogeneity Q-stat to identify traits with high heterogeneity.
- This would indicate that associations between the index genotype and partner’s phenotype may not only act indirectly through causal relationship between the traits, but some direct effect is present. 
- Ideally, for the majority of the traits, the heterogeneity would be low, confirming that (despite what Tenesa suggests) there are no direct index genome to partner phenome effects.
- We tested heterogeneity statistics using the function [mr_heterogeneity](https://mrcieu.github.io/TwoSampleMR/reference/mr_heterogeneity.html) from the [TwoSampleMR](https://mrcieu.github.io/TwoSampleMR/index.html) package.

```{r, het_stots, include = FALSE}
sig_het <- household_MR_summary_AM %>% filter(Het_IVW_Qpval_round < 0.05/num_tests_by_PCs)
```

The number of traits with significant heterogeneity in the MR is: `r dim(sig_het)[1]`.

# Impact of confounders on couple correlation.

We tested the impact of confounding on genetic PCs, N/E coordinates and the following 4 traits:

```{r confounder_traits, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
confounder_trait_tibble <- as_tibble(confounder_traits)
confounder_trait_tibble_descript <- replace_Neale_ID(confounder_trait_tibble, traits_corr2_filled, "value", "outcome_description")
confounder_trait_tibble_descript %>% 
  kbl(col.names = NULL) %>%
  kable_styling()
```

## Ratio of confounding to correlation.

We tested the ratio of confounding to correlation for each confounder trait tested.

```{r corr_due_to_confounding_ratio, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}

corr_impact_by_coords_ratio <- corr_impact_by_coords %>% 
  dplyr::select(Neale_pheno_ID, trait_couple_corr, trait_couple_corr_se, corr_due_to_confounding_all, corr_due_to_confounding_all_se) %>%
  unique() %>%
  mutate(trait_interest = "household_coords") %>%
  mutate(trait_ratio = corr_due_to_confounding_all/trait_couple_corr)

corr_impact_by_PCs_ratio <- corr_impact_by_PCs %>% 
  dplyr::select(Neale_pheno_ID, trait_couple_corr, trait_couple_corr_se, corr_due_to_confounding_all, corr_due_to_confounding_all_se) %>%
  unique() %>%
  mutate(trait_interest = "PCs") %>%
  mutate(trait_ratio = corr_due_to_confounding_all/trait_couple_corr)

corr_impact_by_traits_ratio <- corr_impact_by_traits %>% group_by(trait_interest) %>% 
  dplyr::select(trait_j, trait_interest, trait_j_couple_corr, trait_j_couple_corr_se, 
                corr_due_to_confounding, corr_due_to_confounding_se) %>% 
  mutate(trait_ratio = corr_due_to_confounding/trait_j_couple_corr)

corr_impact_by_traits_ratio <- replace_Neale_ID(corr_impact_by_traits_ratio, traits_corr2_filled, "trait_interest", "outcome_description") %>%
  dplyr::rename(trait_interest = outcome_description)

corr_impact_ratios <- rbind(corr_impact_by_PCs_ratio, corr_impact_by_coords_ratio, corr_impact_by_traits_ratio)
corr_impact_ratios_summary <- corr_impact_ratios %>%
  group_by(trait_interest) %>%
  dplyr::summarise(mean_ratio = mean(trait_ratio), median_ratio = median(trait_ratio), sd_ratio = sd(trait_ratio), number_of_traits = n())

corr_impact_ratios_summary %>%
  kbl() %>%
  kable_styling()

```

*A note on the number of traits:*

We tested the impact of `trait_interest` on couple correlation amongst all traits in our couple MR framework (number of traits = `r dim(household_MR_summary_AM)[1]`) with the exception of home location and birth place coordinates (2 each [N and E] * 2), we also removed the trait itself from the confounder analysis where applicable (in the case of traits `20016_irnt`, `738` and `845`). [Trait `189_irnt` was not included in our final couple MR analysis but could still be evaluated here because no MR-estimates were used.]

## Plots.

The figure below shows the 4 trait which showed the greatest impact on couple correlation (mean confounding ratio > 0.01).

```{r confounder_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}

geography_plots <- list()

plots <- list()
for(i in 1:length(confounder_traits)){

  trait <- confounder_traits[i]
  trait_description <- confounder_trait_tibble_descript[[1]][i]

  if(trait=="738"){trait_description <- "Average total household \nincome before tax"}
  if(trait=="189_irnt"){trait_description <- "Townsend deprivation index \nat recruitment"}

  plot_data <- corr_impact_by_traits %>% filter(trait_interest==trait) %>% filter(trait_interest!=trait_j)
  plot_data <- replace_Neale_ID(plot_data, traits_corr2_filled, "trait_j", "Phenotype")
  plot <-
    ggplot2::ggplot(
      data = plot_data,
      ggplot2::aes(x = trait_j_couple_corr,
                   y = corr_due_to_confounding, Phenotype = Phenotype)
    ) +
    theme(plot.title = element_text(hjust = 0.5)) +
    geom_point(alpha = 3 / 4) +

    ggplot2::geom_errorbar(
      ggplot2::aes(ymin = corr_due_to_confounding - corr_due_to_confounding_se, ymax = corr_due_to_confounding + corr_due_to_confounding_se),
      colour = "grey",
      width = 0
    ) +
    ggplot2::geom_errorbar(
      ggplot2::aes(xmin = trait_j_couple_corr - trait_j_couple_corr_se, xmax = trait_j_couple_corr + trait_j_couple_corr_se),
      colour = "grey",
      width = 0
    ) +

    theme_minimal() +
    ggplot2::labs(
      x = paste0("Couple trait correlation"),
      y = paste0("Couple correlation \ndue to confounding"),
      title = paste0(trait_description)
      #title = paste0("Impact of ", trait_description, "\n on couple correlation")
    ) +
    theme(plot.title = element_text(size=11, hjust = 0.5)) + theme(axis.title = element_text(size = 10))

  plots[[i]] <- plot 
}

figures_grid <- plot_grid(plotlist = plots, nrow=2, ncol = 2, byrow = F, labels="AUTO")
figures_grid
```

# Comparison of paths from index to partner.

As a recap, the model is below.

```{r YiYp_adj_model, echo = FALSE}
library(knitr)
knitr::include_graphics("assets/two_trait_model_yiyp_adj.png", error = FALSE)
```

We sought to compare the different paths from $X_i$ to $Y_p$ using the estimates computed as $\omega$, $\rho$ and $\gamma$.

```{r proxyMR_figure2, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600}
proxyMR_prod_comparison_fig_yiyp_adj
```

We considered the regression of the following estimates vs $\omega$ with an intercept passing through the origin (i.e. $estimate \sim \omega + 0$:

1. $\rho$
2. $\gamma$
3. $\rho$ + $\gamma$
4. $\rho_{resid} + \gamma$, where $\rho$  is residualized for $\gamma$ (i.e. $\rho_{resid} = resid(lm(\rho \sim \gamma))$)

The results are summarized below:

```{r regression_summary, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

data <- proxyMR_comparison_summary_yiyp_adj_SNPmeta
data$rho_resid <- resid(lm(rho_beta ~ gam_beta, data = data))
data$gam_rho_resid <- data$rho_resid + data$gam_beta

result <- numeric()
for(outcome in c("rho", "gam", "gam_rho", "gam_rho_resid")){

  col_to_rename <- paste0(outcome, "_beta")
  if(outcome=="gam_rho_resid") {col_to_rename <- outcome}
  data_model <- data %>% dplyr::rename(outcome = !!col_to_rename)
  model <- summary(lm(outcome ~ omega_beta + 0, data = data_model)) 
  r2 <- model %>% glance %>% pull(r.squared)
  model_summary <- model %>% tidy() %>% 
    mutate(term = "omega") %>% mutate(outcome = outcome) %>% mutate(r.squared = r2)

  result <- rbind(result, model_summary)

}

result %>% dplyr::select(outcome, everything()) %>%
  kbl() %>%
  kable_styling()

```

```{r remove_opp_sign, include = FALSE}

opp_sign_rows <- numeric()
for(col1 in c("gam_beta", "rho_beta", "gam_rho_beta", "gam_rho_resid", "omega_beta")){

  for(col2 in c("gam_beta", "rho_beta", "gam_rho_beta", "gam_rho_resid", "omega_beta")){

      temp <- which(sign(data[[col1]])!=sign(data[[col2]]))
      opp_sign_rows <- c(opp_sign_rows, temp)
  }

}

opp_sign_rows <- unique(opp_sign_rows)

```

To further explore these estimates, performed paired t-tests to find the significant differences between pairs of estimates. We first removed all instances where the sign did not match between any two pairings. In total, of the `r dim(data)[1]` trait pairs under consideration, `r length(opp_sign_rows)` were removed.

```{r boxplot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}

df <- data[-opp_sign_rows,] %>% dplyr::select(exposure_ID, outcome_ID, gam_beta, rho_beta, gam_rho_beta, gam_rho_resid, omega_beta)

data_boxplot  <- df %>%
  pivot_longer(cols = !c("exposure_ID", "outcome_ID"), names_to = "data", values_to = "beta") %>%
  mutate_at("data", as.factor) %>%
  mutate(
    data = fct_recode(as.factor(data),
                      Gamma = "gam_beta",
                      Rho = "rho_beta",
                      Gamma_Rho = "gam_rho_beta",
                      Gamma_Rho_resid = "gam_rho_resid",
                      Omega = "omega_beta")
  )

pairwise.t.test(data_boxplot$beta, data_boxplot$data, p.adjust = "none")

cmpr <- list(c("Omega", "Gamma_Rho"), c("Gamma","Gamma_Rho"), c("Gamma_Rho","Gamma_Rho_resid"), c("Rho", "Gamma_Rho"))

position <- data_boxplot %>% group_by(data) %>% dplyr::summarise(median=median(beta)) %>% 
  arrange(-desc(median)) %>% pull(data)

data_boxplot$data <- factor(data_boxplot$data,levels=position)

boxplot <- data_boxplot %>%
  ggplot( aes(x=beta, y=data)) +
  geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6, option="A") +
  theme_ipsum() +
  theme(
    legend.position="none",
    plot.title = element_text(size=11)
  ) +
  labs(x ="Estimate", y = "") +
  scale_y_discrete(labels=c("Gamma_Rho" = "\u03B3 + \u03C1", "Omega" = "\u03C9",
                            "Gamma" = "\u03B3", "Gamma_Rho_resid" = expression( paste(rho ["resid"], " + ", gamma)), "Rho" = "\u03C1")) +
  stat_compare_means(comparisons = cmpr, tip.length=0.01,
                     label = "p.signif",
                     symnum.args = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1),
                                        symbols = c("****", "***", "**", "*", "ns")))

boxplot

```