### 2022-03-07 analysis/manuscript.Rmd
wflow_publish("analysis/manuscript.Rmd")

---
title: "Manuscript"
author: "Jenny Sjaarda"
date: "2022-01-04"
output: workflowr::wflow_html
---

```{r setup, include = FALSE}

options(scipen=999)
source("code/settings.R")
source("R/functions.R")
library(targets)
library(kableExtra)
library(DT)
library(knitr)
library(ggplot2)
library(kableExtra)
library(cowplot)
library(plotly)
library(plyr)
library(dplyr)
library(ggrepel)
library(viridis)
library(hrbrthemes)
library(ggpubr)
library(broom)
library(data.table)
library(xlsx)
```

```{r load_targets, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

tar_load(couple_MR_vs_trait_corr)
tar_load(corr_impact_by_PCs)
tar_load(corr_impact_by_traits)
tar_load(corr_impact_by_coords)
tar_load(corr_potential_trait_confounders)
tar_load(traits_corr)
tar_load(PCs_corr)
tar_load(traits_corr2_filled)
tar_load(household_MR_summary_AM)
tar_load(num_tests_by_PCs)
tar_load(num_tests_by_PCs_AM_sig)
tar_load(household_MR_binned_het_sex_spec)
tar_load(household_MR_binned_het_SNPmeta)
tar_load(traits_final)
tar_load(household_MR_AM_FvsM_fig)
tar_load(confounder_traits)
tar_load(proxyMR_prod_comparison_fig_yiyp_adj)
tar_load(proxyMR_comparison_summary_yiyp_adj_SNPmeta)
tar_load(household_MR_summary_AM)
tar_load(household_MR_binned_het_SNPmeta)
tar_load(potential_trait_confounders_MVMR)
tar_load(corr_mat_traits)

```

# A few notes (for myself).

- The number of effective tests can be found at the target `num_tests_by_PCs`. 
- The number of effective tests among those significant in the MR can be found at the target `num_tests_by_PCs_AM_sig`.

# Pearson vs. Spearman correlation.

We wanted to ensure that INQT was not significantly impacting the correaltions of each trait, so we also calculated a non-parametric based correaltion (Spearman) and found consistent correlations.

```{r pearson_vs_spearman, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE, warning = FALSE}
tar_load(Neale_SGG_dir_filt)
tar_load(traits_corr_spearman)
tar_load(traits_corr)

Neale_SGG_dir_filt_unique <- Neale_SGG_dir_filt[,c("phenotype", "phenotype_sub", "sgg_phesant_name", "description", "SGG_available", "variable_type")] %>% unique()

pearson_vs_spearman_dat <- inner_join(traits_corr, unique(Neale_SGG_dir_filt_unique[,c("sgg_phesant_name", "variable_type")]), by = c("ID"="sgg_phesant_name")) %>% dplyr::rename(r2_pearson = r2) %>%
  inner_join(traits_corr_spearman[,c("ID", "r2")]) %>% dplyr::rename(r2_spearman = r2) %>% 
  as_tibble() %>% type_convert() %>%
  mutate(r_pearson = sqrt(r2_pearson)) %>% 
  mutate(r_spearman = sqrt(r2_spearman))

pearson_vs_spearman_plot <- ggplot2::ggplot(data = pearson_vs_spearman_dat, 
                                            ggplot2::aes(x = r_pearson,
                                                         y = r_spearman, 
                                                         color = factor(variable_type))) + 
  geom_point() +
  theme_minimal() +
  ggplot2::labs(x = paste0("Pearson correlation"),
                y = paste("Spearman correlation")) + 
  scale_color_manual(name = "Variable type", values = custom_col[c(1,2,3)], labels=c("Binary","Continuous (IRNT)", "Ordinal"))

pearson_vs_spearman_plot
```

# Relationship between causal effects and raw phenotypic correlation in couples.

Sought to compare the raw correlation amongst couples vs. the standardized MR effects. There are many traits where the `correlation > MR effects` including some traits of interest such as standing height, place of birth, among many others.

For example, if we take a trait such as place of birth (NC), this is very highly correlated among couples. However, the within-couple MR-estimate is smaller than the observed phenotypic correaltion. This is the case for many such traits.

```{r pheno_corr_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE, warning = FALSE}

### phenotypic correlation plot for 129_irnt (place of birth)

tar_load(path_pheno_data)
tar_load(hh_pairs_filter)
Neale_pheno_ID <- "129_irnt" ## this is the Neale_id, used to be pheno_description
phes_ID <- gsub("_irnt", "", Neale_pheno_ID)

male_pheno_file <- path_pheno_data[which(endsWith(path_pheno_data, paste0("/phesant/", Neale_pheno_ID, "_male.txt")))]
female_pheno_file <- path_pheno_data[which(endsWith(path_pheno_data, paste0("/phesant/", Neale_pheno_ID, "_female.txt")))]
male_pheno_data <- fread(male_pheno_file, header=TRUE, select=c("IID","sex","age", phes_ID))
female_pheno_data <- fread(female_pheno_file, header=TRUE, select=c("IID","sex","age", phes_ID))

joint_data <- rbind(male_pheno_data, female_pheno_data)

pairs_filter_copy <- hh_pairs_filter
    pairs_filter_copy <- pairs_filter_copy %>%
      mutate(HOUSEHOLD_MEMBER1_M = case_when(HOUSEHOLD_MEMBER1_sex == 1 ~ HOUSEHOLD_MEMBER1,
                                             HOUSEHOLD_MEMBER2_sex == 1 ~ HOUSEHOLD_MEMBER2)) %>%

      mutate(HOUSEHOLD_MEMBER2_F = case_when(HOUSEHOLD_MEMBER1_sex == 0 ~ HOUSEHOLD_MEMBER1,
                                             HOUSEHOLD_MEMBER2_sex == 0 ~ HOUSEHOLD_MEMBER2)) %>%

      dplyr::select(-HOUSEHOLD_MEMBER1, -HOUSEHOLD_MEMBER2) %>%
      dplyr::rename(HOUSEHOLD_MEMBER1 = HOUSEHOLD_MEMBER1_M) %>%
      dplyr::rename(HOUSEHOLD_MEMBER2 = HOUSEHOLD_MEMBER2_F)

pairs_filter_copy$trait1 <- joint_data[[phes_ID]][match(pairs_filter_copy[["HOUSEHOLD_MEMBER1"]], joint_data$IID)]
pairs_filter_copy$trait2 <- joint_data[[phes_ID]][match(pairs_filter_copy[["HOUSEHOLD_MEMBER2"]], joint_data$IID)]

pheno_corr_dat <- pairs_filter_copy

pheno_corr_plot <- ggplot2::ggplot(data = pheno_corr_dat, ggplot2::aes(x = trait1,
                                                       y = trait2)) +
  geom_point(alpha = 1.2, color = "gray") +
  #geom_text(show.legend = FALSE) +
  #geom_abline(slope=1, intercept=0) +

  theme_minimal() +

  #geom_text(aes(x=1,y=2,label=""),show.legend=F) +
  ggplot2::labs(x = paste0("Male place of birth (NC)"),
                y = paste("Female place of birth (NC)")) + 
  geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) +
  theme_ipsum()

pheno_corr_plot

```

A plot comparing the two effects, and a summary table of the significant effects is shown below.

```{r corr_vs_MR_effects, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}

interesting_phenotypes <- c("Basal metabolic rate", "Place of birth in UK - north co-ordinate", 
                            "Systolic blood pressure, automated reading", "Standing height", 
                            "Weight", "Forced vital capacity (FVC), Best measure",
                            "Basophill count", "Pulse rate, automated reading")

corr_vs_MR_effects_dat <- replace_Neale_ID(couple_MR_vs_trait_corr, traits_corr2_filled, "exposure_ID", "outcome_description") %>% 
  mutate(Phenotype = outcome_description) %>% 
  mutate(Phenotype = case_when(Phenotype == "Basal metabolic rate" ~ "BMR",
                               Phenotype == "Systolic blood pressure, automated reading" ~ "SBP",
                               Phenotype == "Pulse rate, automated reading" ~ "Pulse rate",
                               Phenotype == "Body fat percentage" ~ "BFP",
                               Phenotype == "Standing height" ~ "Height",
                               Phenotype == "Reticulocyte percentage" ~ "Reticulocyte %",

                               #Phenotype == "Weight" & outcome_ID == "21002_irnt" ~ "Weight",
                               Phenotype == "Forced vital capacity (FVC), Best measure" ~ "FVC",
                               Phenotype == "Red blood cell (erythrocyte) count" ~ "RBC count",
                               Phenotype == "Place of birth in UK - north co-ordinate" ~ "Place of birth, NC",
                               TRUE ~ Phenotype)) %>%

  dplyr::rename(Correlation = couple_r) %>% dplyr::rename(MR_est = IVW_beta) %>% 
  mutate(diff_p_sig = case_when(diff_p < 0.05/num_tests_by_PCs ~ TRUE, TRUE ~ FALSE)) %>%
  mutate(outcome_description2 = case_when(outcome_description=="Weight" & outcome_ID == "23098_irnt" ~ "Weight2",
                                          TRUE ~ outcome_description)) %>%
  mutate(label_phenos = case_when(outcome_description2 %in% interesting_phenotypes ~ TRUE, TRUE ~ FALSE))

  # TRUE - BF sig difference 
  # FALSE - non-sig 
  #mutate(diff_p_sig = ifelse(diff_p < 0.05/num_tests_by_PCs, TRUE, FALSE))

corr_vs_MR_effects_plot <- ggplot2::ggplot(data = corr_vs_MR_effects_dat, ggplot2::aes(x = Correlation,
                                                       y = MR_est, 
                                                       label = Phenotype, 
                                                       color = factor(diff_p_sig)),) +
  geom_point(alpha = 1.2) +
  scale_color_manual(values = custom_col[c(1,2)], labels=c("Non-significant","BF-signifncant difference")) +
  #geom_text(show.legend = FALSE) +
  geom_abline(slope=1, intercept=0) +

  ggplot2::geom_errorbar(
    ggplot2::aes(ymin = MR_est - IVW_se, ymax = MR_est + IVW_se),
    colour = "grey",
    width = 0
  ) +
  ggplot2::geom_errorbar(
    ggplot2::aes(xmin = Correlation - couple_r_se, xmax = Correlation + couple_r_se),
    colour = "grey",
    width = 0
  ) +

  theme_minimal() +
  theme(legend.text=element_text(size=11)) +
  #geom_text(aes(x=1,y=2,label=""),show.legend=F) +
  ggplot2::labs(x = paste0("Couple correlation"),
                y = paste("Standardized MR effect")) +
  theme(legend.position="bottom") + theme(legend.title=element_blank()) +
  geom_text_repel(data = subset(corr_vs_MR_effects_dat, label_phenos==TRUE),
                  nudge_y = -0.25 - subset(corr_vs_MR_effects_dat, label_phenos==TRUE)$MR_est, size = 3.25,
                  direction = "x", show.legend = FALSE )

corr_vs_MR_effects_plot

corr_vs_MR_effects_sig_dat <- corr_vs_MR_effects_dat %>% dplyr::rename(IVW_beta = MR_est) %>% 
  filter(diff_p < 0.05/num_tests_by_PCs) %>%
  mutate(mr_corr_diff = Correlation - IVW_beta) %>% 
  mutate(correlation_larger = ifelse(Correlation > IVW_beta, TRUE, FALSE)) %>% 
  arrange(diff_p) %>% dplyr::select(-exposure_ID, -Phenotype, -outcome_description2, -label_phenos, -mr_corr_diff, -diff_p_sig) %>% 
  dplyr::rename(trait_description = outcome_description) %>%
  dplyr::rename(trait_ID = outcome_ID) %>%
  dplyr::rename(couple_r = Correlation) %>%
  dplyr::select(trait_description, trait_ID, everything())

## prune for correlations < 0.8

corr_vs_MR_effects_sig_dat.DT <- corr_vs_MR_effects_sig_dat %>% 
  dplyr::select(trait_description, trait_ID, couple_r, IVW_beta, IVW_pval, diff_p, correlation_larger)

#prune_pheno_table(corr_vs_MR_effects_sig_dat, "trait_ID", corr_mat_traits, corr_trait_threshold)

# Could prune the results, but only one trait is removed so seems pointless:

# Result are pruned to only show traits with correlation < `r corr_trait_threshold`. (The number of traits before pruning was `r dim(corr_vs_MR_effects_sig_dat)[1]` vs `r dim(corr_vs_MR_effects_sig_dat_pruned.DT)[1]` after pruning).
```

## Summary of cases where correlation is different from the MR estimate.

The table below displays the `r dim(corr_vs_MR_effects_sig_dat.DT)[1]` traits that have correlation significantly different from the MR-estimate. Of the significant difference `r table(corr_vs_MR_effects_sig_dat$correlation_larger)[[1]]` have MR-estimates larger than than correlation, suggesting the presence of a negative confounder.

```{r corr_vs_MR_effects_table, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}

format_round_cols <- colnames(dplyr::select_if(corr_vs_MR_effects_sig_dat.DT, is.numeric))

DT::datatable(corr_vs_MR_effects_sig_dat.DT %>% dplyr::select(-trait_ID), width = "100%") %>% formatSignif(columns=format_round_cols, digits = 3)

```

## Search for potential confounders to explain cases where correlation differs from MR estimate.

1. Restrict to only traits that have correlation significantly different from MR-estimate (`r dim(corr_vs_MR_effects_sig_dat)[1]` traits have correlation different from MR_estimate).
2. For each of these traits, find the traits which have a causal effect on this trait (MR p-val < 0.05/[number of effective tests]).
3. Calculate the correlation due to confounding as: $C = (\alpha_{z\rightarrow x})^2 * \alpha_{z_i\rightarrow z_p}$.
4. Calculate the ratio of this correlation ($C$) and the correlation of $X$ in partners ($C/r_x$). 
5. Remove any `potential_counders` with correlation with `trait_ID` > `r corr_trait_threshold`.
6. For each `trait_ID` prune remaining `potential_counders` so that all have correlations <  `r corr_trait_threshold`, prioritized by their `corr_due_to_confounding_ratio`.

A summary of the top 5 confounders for each `trait_ID` can be seen below:

```{r simple_C_model, echo = FALSE}
library(knitr)
knitr::include_graphics("assets/simple_confounder_model.png", error = FALSE)

```

```{r confounder_summary, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}

number_of_potential_confounders <- dim(corr_potential_trait_confounders)[1]/dim(corr_vs_MR_effects_sig_dat)[1]

non_sig_confounders <- corr_potential_trait_confounders %>% filter(!sig_confounder) %>% group_by(outcome_ID) %>%
  dplyr::summarize(description = outcome_description[1], n = n()) %>% 
  filter(n==number_of_potential_confounders) %>%
  mutate(n = n-number_of_potential_confounders) %>% mutate(potential_confounders = NA) %>%
  mutate(mean_C_ratio = NA) %>%
  dplyr::rename(trait_ID = outcome_ID)

sig_confounders <- corr_potential_trait_confounders %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
  arrange(-corr_due_to_confounding_ratio) %>%
  mutate(total_potential_confounders = n()) %>%
  dplyr::summarize(description = outcome_description[1],
                   total_potential_confounders = total_potential_confounders[1],
                   potential_confounders = paste0(exposure_description, collapse = "; "), 
                   max_C = max(corr_due_to_confounding), 
                   min_C = min(corr_due_to_confounding)) %>%
  dplyr::rename(trait_ID = outcome_ID) %>%
  dplyr::rename(trait_description = description) %>% 
  dplyr::select(trait_description, trait_ID, everything())

sig_confounders_top5 <- corr_potential_trait_confounders %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
  arrange(-corr_due_to_confounding_ratio) %>%
  mutate(total_potential_confounders = n()) %>%
  slice_head(n = 5) %>%
  #mutate(potential_confounders = paste0(exposure_description, collapse = "; ")) %>%
  dplyr::summarize(trait_description = outcome_description[1],
                   total_potential_confounders = total_potential_confounders[1],
                   top5_potential_confounders = paste0(exposure_description, collapse = "; "), 
                   max_C = max(corr_due_to_confounding), 
                   min_C = min(corr_due_to_confounding)) %>%
  dplyr::rename(trait_ID = outcome_ID) %>%
  left_join(corr_vs_MR_effects_sig_dat.DT, by = c("trait_description", "trait_ID")) %>%
  dplyr::select(trait_description, trait_ID, couple_r, IVW_beta, IVW_pval,everything())

format_round_cols <- colnames(dplyr::select_if(sig_confounders_top5, is.numeric))

DT::datatable(sig_confounders_top5, width = "100%") %>% formatSignif(columns=format_round_cols, digits = 3)

```

We compared the difference in phenotypic estimate and MR-estimates to the maximum C for each trait, as shown in the figure below:

```{r max_C_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}

max_C_data <- corr_potential_trait_confounders %>% filter(sig_confounder) %>% group_by(outcome_ID) %>%
  arrange(-corr_due_to_confounding_ratio) %>%
  mutate(total_potential_confounders = n()) %>%
  slice_head(n = 1) %>%
  left_join(couple_MR_vs_trait_corr %>% 
              dplyr::rename(outcome_ID_IVW_beta = IVW_beta) %>%
              dplyr::rename(outcome_ID_IVW_se = IVW_se) %>%
              dplyr::rename(outcome_ID_couple_r = couple_r) %>% 
              dplyr::rename(outcome_ID_couple_r_se = couple_r_se) %>% 
              dplyr::select(outcome_ID, outcome_ID_IVW_beta, outcome_ID_IVW_se, outcome_ID_couple_r, outcome_ID_couple_r_se)) %>%
  mutate(same_trait_corr_minus_MR_est = abs(outcome_ID_couple_r)- abs(outcome_ID_IVW_beta)) %>%

  dplyr::rename(Z_beta = IVW_beta) %>%
  dplyr::rename(Z_se = IVW_se) %>%
  dplyr::rename(Z_pval = IVW_pval) %>%
  mutate(Z_sq_beta = Z_beta^2) %>%
  mutate(Z_sq_se = sqrt(4*Z_beta^2*Z_se^2+2*Z_se^4)) %>%
  dplyr::rename(Z_AM_beta = exposure_ID_AM_IVW_beta) %>%
  dplyr::rename(Z_AM_se = exposure_ID_AM_IVW_se) %>%
  mutate(Z_prod = Z_sq_beta*Z_AM_beta) %>%
  mutate(Z_prod_var = variance_of_product(Z_sq_beta, Z_sq_se, Z_AM_beta, Z_AM_se)) %>%
  mutate(Z_prod_se = sqrt(Z_prod_var)) %>%
  mutate(same_trait_corr_minus_MR_est_se = variance_of_sum(outcome_ID_IVW_se, outcome_ID_couple_r_se))

max_C_plot <- ggplot2::ggplot(data = max_C_data, ggplot2::aes(x = corr_due_to_confounding,
                                                       y = same_trait_corr_minus_MR_est)) +
  geom_point(alpha = 1.2) +
  #scale_color_manual(values = custom_col[c(1,2)], labels=c("Non-significant","BF-signifncant difference")) +
  #geom_text(show.legend = FALSE) +
  geom_abline(slope=1, intercept=0) +

  ggplot2::geom_errorbar(
    ggplot2::aes(ymin = same_trait_corr_minus_MR_est - same_trait_corr_minus_MR_est_se, ymax = same_trait_corr_minus_MR_est + same_trait_corr_minus_MR_est_se),
    colour = "grey",
    width = 0
  ) +
  ggplot2::geom_errorbar(
    ggplot2::aes(xmin = corr_due_to_confounding - Z_prod_se, xmax = corr_due_to_confounding + Z_prod_se),
    colour = "grey",
    width = 0
  ) +

  theme_minimal() +
  ggplot2::labs(x = paste0("max C"),
                y = paste("abs(correlation) - abs(MR-estimate)")) +
  theme(legend.title=element_blank()) +
  #geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) +
  theme_ipsum()

max_C_plot

```

Finally, we computed a MVMR to sum up the effect of all potential confounders as shown in the DAG below and compared these C-statistics to the difference between phenotypic correlation and MR-estimates.

```{r MVMR_C_model, echo = FALSE}
library(knitr)
knitr::include_graphics("assets/MVMR_confounder_model.png", error = FALSE)
```

```{r MVMR_C_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}

discrepentant_traits <- unique(corr_potential_trait_confounders$outcome_ID)

output <- numeric()
for(i in 1:length(potential_trait_confounders_MVMR)){

  outcome_ID <- discrepentant_traits[i]

  corr_potential_trait_confounders_i <- corr_potential_trait_confounders %>% filter(outcome_ID == !!outcome_ID)

  MVMR_result_org <- potential_trait_confounders_MVMR[[i]]

  MVMR_result_org <- MVMR_result_org %>% mutate(beta_2_var = 4*beta^2*se^2+2*se^4) %>%
    mutate(beta_2_se = sqrt(beta_2_var)) %>% 
    left_join(corr_potential_trait_confounders_i %>% dplyr::select(exposure_ID, exposure_ID_AM_IVW_beta, exposure_ID_AM_IVW_se), by = c("exposure" = "exposure_ID")) %>%
    dplyr::rename(MVMR_Z_beta = beta) %>%
    dplyr::rename(MVMR_Z_se = se) %>%
    dplyr::rename(MVMR_Z_pval = pval) %>%
    mutate(MVMR_Z_sq_beta = MVMR_Z_beta^2) %>%
    dplyr::rename(MVMR_Z_sq_se = beta_2_se) %>% 
    dplyr::rename(Z_AM_beta = exposure_ID_AM_IVW_beta) %>%
    dplyr::rename(Z_AM_se = exposure_ID_AM_IVW_se) %>%
    mutate(Z_prod = MVMR_Z_sq_beta*Z_AM_beta) %>%
    mutate(Z_prod_var = variance_of_product(MVMR_Z_sq_beta, MVMR_Z_sq_se, Z_AM_beta, Z_AM_se))

  MVMR_result_nom <- MVMR_result_org %>% filter(MVMR_Z_pval < 0.05)
  MVMR_result_BF <- MVMR_result_org %>% filter(MVMR_Z_pval < 0.05/dim(MVMR_result_org)[1])

  num_confounders_nom <- dim(MVMR_result_org)[1]
  num_confounders_BF <- dim(MVMR_result_BF)[1]

  row <- cbind(outcome_ID)
  for(filter_type in c("nom", "BF")){

    if(filter_type == "nom"){
      t <- MVMR_result_org %>% filter(MVMR_Z_pval < 0.05)
    }
    if(filter_type == "BF"){
      t <- MVMR_result_org %>% filter(MVMR_Z_pval < 0.05/dim(MVMR_result_org)[1])
    }

    num_confounders <- dim(t)[1]
    C_sum <- sum(t$Z_prod)
    C_sum_var <- sum(t$Z_prod_var)
    C_sum_se <- sqrt(C_sum_var)

    temp_row <- cbind(num_confounders, C_sum, C_sum_var, C_sum_se)
    colnames(temp_row) <- paste0(c("num_confounders", "C_sum", "C_sum_var", "C_sum_se"), "_", filter_type)
    row <- cbind(row, temp_row)

  }

  output <- rbind(output, row)

}

output <- output %>% as_tibble() %>% type_convert() %>% mutate(C_sum_pval = 2*pnorm(-abs(C_sum_nom/C_sum_se_nom)))

MVMR_C_data <- max_C_data %>% left_join(output)

MVMR_C_plot <- ggplot2::ggplot(data = MVMR_C_data, ggplot2::aes(x = C_sum_nom ,
                                                       y =  same_trait_corr_minus_MR_est)) +
  geom_point(alpha = 1.2) +
  #scale_color_manual(values = custom_col[c(1,2)], labels=c("Non-significant","BF-signifncant difference")) +
  #geom_text(show.legend = FALSE) +
  #geom_abline(slope=1, intercept=0) +

  # ggplot2::geom_errorbar(
  #   ggplot2::aes(ymin = MR_est - IVW_se, ymax = MR_est + IVW_se),
  #   colour = "grey",
  #   width = 0
  # ) +
  # ggplot2::geom_errorbar(
  #   ggplot2::aes(xmin = Correlation - couple_r_se, xmax = Correlation + couple_r_se),
  #   colour = "grey",
  #   width = 0
  # ) +

  theme_minimal() +
  ggplot2::labs(x = paste0("MVMR C-sum (p < 0.05)"),
                y = paste("abs(correlation) - abs(MR-estimate)")) +
  theme(legend.title=element_blank()) +
  geom_abline(slope=1, intercept=0)

MVMR_C_plot

```

# Effect of sex, age and time together on causal effects in couples.

```{r AM_data_prep, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE, include = FALSE}

AM_result_sig <- household_MR_summary_AM %>% filter(IVW_pval < 0.05/num_tests_by_PCs) 
AM_sig_traits <-household_MR_summary_AM %>% filter(IVW_pval < 0.05/num_tests_by_PCs) %>% pull(exposure_ID) %>% unique()
AM_sex_spec_result_sig <- household_MR_binned_het_sex_spec %>% filter(same_trait ==TRUE) %>% filter(exposure_ID %in% AM_sig_traits) %>% filter(sex_het_pval < 0.05/num_tests_by_PCs_AM_sig)

```

- For each trait in the pipeline, we tested the causal effect within couples (i.e. the number of analyses corresponds to the number of traits (`r dim(traits_final)[1]`)). 
- The significant results after adjusting for multiple hypothesis testing are shown below. 
- `IVW_meta_beta` and `IVW_meta_pval` corresponds to the beta and p-value of the meta-analyzed MR across sexes, respectively (i.e. MR estimates were computed in each sex-seperately using sex-specific SNP-exposure and SNP-outcome results and then meta-analyzed).  
- After adjusting for multiple tests (p < `0.05/`r num_tests_by_PCs``), identified `r dim(AM_result_sig)[1]` significant assortative mating (same-trait) MR results (corresponding to the table below)..

```{r, AM_sig, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE, include = TRUE}

AM_result_sig.DT <- AM_result_sig %>% dplyr::select(exposure_description, exposure_ID, IVW_beta, IVW_se, IVW_pval) %>% unique() %>% 
  dplyr::rename(trait_description = exposure_description) %>%
  dplyr::rename(trait_ID = exposure_ID)

DT::datatable(AM_result_sig.DT, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_se', 'IVW_pval'), digits=3)

AM_sig_sex_het_nom <- household_MR_binned_het_sex_spec %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% filter(sex_het_pval < 0.05)
AM_sig_sex_het_nom <- replace_Neale_ID(AM_sig_sex_het_nom, traits_corr2_filled,"exposure_ID", "exposure_description") %>%
  dplyr::rename(trait_description = exposure_description) %>%
  dplyr::rename(trait_ID = exposure_ID) %>%
  dplyr::select(trait_description, trait_ID, everything())

AM_result_sig_nom.DT <- AM_sig_sex_het_nom %>% dplyr::select(trait_description, trait_ID, exposure_sex, IVW_beta, IVW_se, IVW_pval, sex_het_pval) %>% unique()
```

## Sex heterogeneity.

- For each AM MR ($X_{p} \sim X_i$), analyses were run in each sex separately. 
- Of the `r dim(AM_sig_traits)[1]` significant results shown above, we tested to see if there was any difference in AM MR estimates between sexes. 
- A similar adjustment for multiple hypothesis testing was applied as above, resulting in `r num_tests_by_PCs_AM_sig` number of effective tests. 
- After adjusting for multiple hypothesis testing (p < `0.05/`r num_tests_by_PCs_AM_sig``), `r dim(AM_sex_spec_result_sig)[1]` traits showed significant differences amongst sexes.
- The table below shows the *nominally significant* results at `p < 0.05` (of which there are `r dim(AM_result_sig_nom.DT)[1]/2`).

```{r AM_sex_het_nom, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}

DT::datatable(AM_result_sig_nom.DT, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval', 'sex_het_pval'), digits=3)

nominally_sig_sex_differences <- AM_result_sig_nom.DT %>% pivot_wider(names_from = exposure_sex, values_from = c(IVW_beta, IVW_se, IVW_pval)) 
# write.table(nominally_sig_sex_differences, "docs/manuscript_tables/nominally_sig_sex_differences.txt", row.names = F, sep=",", quote = F)
```

## MR binned results (exploring effect of age and time-together).

```{r binned_results_prep, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE, include = FALSE}

slope_sig_age <- household_MR_binned_het_SNPmeta %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
  filter(grouping_var == "age_even_bins") %>% 
  filter(bin_slope_pval < 0.05/num_tests_by_PCs_AM_sig)

slope_sig_tt <- household_MR_binned_het_SNPmeta %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
  filter(grouping_var == "time_together_even_bins") %>% 
  filter(bin_slope_pval < 0.05/num_tests_by_PCs_AM_sig)

slope_wt_sig_age <- household_MR_binned_het_SNPmeta %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
  filter(grouping_var == "age_even_bins") %>% 
  filter(bin_slope_pval_wt < 0.05/num_tests_by_PCs_AM_sig)

slope_wt_sig_tt <- household_MR_binned_het_SNPmeta %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
  filter(grouping_var == "time_together_even_bins") %>% 
  filter(bin_slope_pval_wt < 0.05/num_tests_by_PCs_AM_sig)

sex_het_sig_age <- household_MR_binned_het_sex_spec %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
  filter(grouping_var == "age_even_bins") %>% 
  filter(bin_slope_sex_het_pval < 0.05/num_tests_by_PCs_AM_sig)

sex_het_sig_tt <- household_MR_binned_het_sex_spec %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
  filter(grouping_var == "time_together_even_bins") %>% 
  filter(bin_slope_sex_het_pval < 0.05/num_tests_by_PCs_AM_sig)

sex_het_wt_sig_age <- household_MR_binned_het_sex_spec %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
  filter(grouping_var == "age_even_bins") %>% 
  filter(bin_slope_sex_het_pval_wt < 0.05/num_tests_by_PCs_AM_sig)

sex_het_wt_sig_tt <- household_MR_binned_het_sex_spec %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% 
  filter(grouping_var == "time_together_even_bins") %>% 
  filter(bin_slope_sex_het_pval_wt < 0.05/num_tests_by_PCs_AM_sig)

```

- For each AM MR ($X_{p} \sim X_i$), analyses were run in the full sample as well as in 5 roughly equal sized bins according to time couple had been together (`time_together_even_bins`) estimated using [time at household variable](https://biobank.ctsu.ox.ac.uk/showcase/field.cgi?id=699), and median age (`age_even_bins`). 
- Of the `r length(AM_sig_traits)` significant results shown above, we tested to see if there was any significant difference in AM MR estimates amongst the two grouping variables. 
- Specifically, the outcome data set was split into 5 bins, and the SNP-outcome effect was estimated in each bin separately (and each sex seperately). 
- These outcome-SNP effects were then used to generate bin-specific MR estimates, using the same SNP-exposure effects from Neale. 
- To estimate the difference among bins, we used two approaches: 
    1. The Cochran's Q test to test for heterogeneity in meta-analyses *(of which we obtained no significant results, so not presenting in paper)*.
    2. Tested the significance of the slope of linear model of median bin (either `age` or `time together`) versus the bin-specific MR estimate.

------------------------------------------------------------------------

Slopes were calculated by estimating the beta-coefficient of a linear regression between MR estimate within each bin (dependent variable) and median bin (independent variable, either age or time at same household). Linear models were run both *unweighted* and *weighted* for the by the inverse of the SE of the MR estimate.

The number of significant trends ($\beta$ estimates from the model: $\alpha_{bin} \sim median_{bin}$) significant after multiple hypothesis testing (p < `0.05/`r num_tests_by_PCs_AM_sig` = `r 0.05/num_tests_by_PCs_AM_sig``) in each group was:

  - Binned by median *age* of couples, number of significant results after adjusting for number of tests:
      - $\beta$ unweighted, meta-anlayzed across sexes: `r dim(slope_sig_age)[1]`
      - $\beta$ weighted by inverse of the SE, meta-anlayzed across sexes: `r dim(slope_wt_sig_age)[1]`
      - $\beta$ unweighted, significantly different between sexes: `r dim(sex_het_sig_age)[1]/2`
      - $\beta$ weighted by inverse of the SE, significantly different between sexes: `r dim(sex_het_wt_sig_age)[1]/2`

  - Binned by *time* couple had been together, number of significant results after adjusting for number of tests:
      - $\beta$ unweighted, meta-anlayzed across sexes: `r dim(slope_sig_tt)[1]`
      - $\beta$ weighted by inverse of the SE, meta-anlayzed across sexes: `r dim(slope_wt_sig_tt)[1]`
      - $\beta$ unweighted, significantly different between sexes: `r dim(sex_het_sig_tt)[1]/2`.
      - $\beta$ weighted by inverse of the SE, significantly different between sexes: `r dim(sex_het_wt_sig_tt)[1]/2`

The significant traits correspond to the following traits:

```{r sig_sex_diff, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE, include = TRUE}

sex_het_wt_sig_age_traits <- replace_Neale_ID(sex_het_wt_sig_age, traits_corr2_filled, "exposure_ID", "outcome_description") %>% pull(outcome_description) %>% unique()

sex_het_sig_age_sub <- sex_het_sig_age %>% dplyr::select(exposure_ID, exposure_sex, grouping_var, bin_slope_beta, bin_slope_se, bin_slope_pval, bin_slope_sex_het_pval)
sex_het_sig_tt_sub <- sex_het_sig_tt %>% dplyr::select(exposure_ID, exposure_sex, grouping_var, bin_slope_beta, bin_slope_se, bin_slope_pval, bin_slope_sex_het_pval)

sig_sex_diff_dat <- replace_Neale_ID(rbind(sex_het_sig_age_sub, sex_het_sig_tt_sub), traits_corr2_filled, "exposure_ID", "exposure_description")

DT::datatable(sig_sex_diff_dat, width = "100%") %>% formatSignif(columns=c('bin_slope_beta', 'bin_slope_se', 'bin_slope_pval', 'bin_slope_sex_het_pval'), digits=3)

```

A few potential figures we could include for a given trait are shown below.

```{r binned_figures, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE, include = TRUE}

tar_load(exposure_info)

branches_to_pull <- numeric()
for(descriptions in unique(sig_sex_diff_dat$exposure_description))
{

  for(i in 1:length(exposure_info)){
  temp <- exposure_info[[i]]
  if(temp[which(temp$Value=="description"), "Info"] == descriptions) {break}
  }
  branches_to_pull <- c(i, branches_to_pull)
}

figs <- list()

for(i in branches_to_pull){

  name <- paste0("exposure_", i)
  figs[[name]] <- tar_read(household_MR_binned_AM_figs, branches = i)[[1]]

}

figs[[1]][[4]]
figs[[1]][[5]]
figs[[2]][[4]]
figs[[2]][[5]]

```

## Phenotypic binned results (exploring effect of age and time-together).

We also tested the phenotypic correlation among the different bins. Below are the traits that show a significant trend either with Pearson's or Spearman's correlation (p < `r 0.05/num_tests_by_PCs`).

```{r, pheno_binned_result, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}

tar_load(binned_pheno_corrs)

binned_pheno_corr_summary <- bind_rows(lapply(binned_pheno_corrs, function(x) {x[[2]]}))

binned_pheno_corr_sig <- binned_pheno_corr_summary %>% filter(bin_slope_pval_spearman < 0.05/num_tests_by_PCs | bin_slope_pval_pearson < 0.05/num_tests_by_PCs)

binned_pheno_corr_sig.DT <- replace_Neale_ID(binned_pheno_corr_sig, traits_corr2_filled, "trait", "trait_description") %>%
  dplyr::rename(trait_ID = trait) %>%
  dplyr::select(trait_description, trait_ID, everything())

format_round_cols <- colnames(dplyr::select_if(binned_pheno_corr_sig.DT, is.numeric))

DT::datatable(binned_pheno_corr_sig.DT, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3)

# write.table(binned_pheno_corr_sig.DT, "docs/manuscript_tables/binned_phenotypic_correlations.txt", row.names = F, sep=",", quote = F)

```

# Heterogeneity stats.

- In the single trait analysis, we also decided to investigate Cochran’s heterogeneity Q-stat to identify traits with high heterogeneity.
- This would indicate that associations between the index genotype and partner’s phenotype may not only act indirectly through causal relationship between the traits, but some direct effect is present. 
- Ideally, for the majority of the traits, the heterogeneity would be low, confirming that (despite what Tenesa suggests) there are no direct index genome to partner phenome effects.
- We tested heterogeneity statistics using the function [mr_heterogeneity](https://mrcieu.github.io/TwoSampleMR/reference/mr_heterogeneity.html) from the [TwoSampleMR](https://mrcieu.github.io/TwoSampleMR/index.html) package.

```{r, het_stats, include = FALSE}
sig_het <- household_MR_summary_AM %>% filter(Het_IVW_Qpval_round < 0.05/num_tests_by_PCs)
```

The number of traits with significant heterogeneity in the MR is: `r dim(sig_het)[1]`.

# Impact of confounders on couple correlation.

We tested the impact of confounding on genetic PCs, N/E coordinates and the following 4 traits:

```{r confounder_traits, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}
confounder_trait_tibble <- as_tibble(confounder_traits)
confounder_trait_tibble_descript <- replace_Neale_ID(confounder_trait_tibble, traits_corr2_filled, "value", "outcome_description")
confounder_trait_tibble_descript %>% 
  kbl(col.names = NULL) %>%
  kable_styling()
```

## Ratio of confounding to correlation.

We tested the ratio of confounding to correlation for each confounder trait tested.

```{r corr_due_to_confounding_ratio, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}

corr_impact_by_coords_ratio <- corr_impact_by_coords %>% 
  dplyr::select(Neale_pheno_ID, trait_couple_corr, trait_couple_corr_se, corr_due_to_confounding_all, corr_due_to_confounding_all_se) %>%
  unique() %>%
  mutate(confounder_trait = "household_coords") %>%
  mutate(trait_ratio = corr_due_to_confounding_all/trait_couple_corr)

corr_impact_by_PCs_ratio <- corr_impact_by_PCs %>% 
  dplyr::select(Neale_pheno_ID, trait_couple_corr, trait_couple_corr_se, corr_due_to_confounding_all, corr_due_to_confounding_all_se) %>%
  unique() %>%
  mutate(confounder_trait = "PCs") %>%
  mutate(trait_ratio = corr_due_to_confounding_all/trait_couple_corr)

corr_impact_by_traits_ratio <- corr_impact_by_traits %>% dplyr::rename(confounder_trait = trait_interest) %>% 
  group_by(confounder_trait) %>% 
  dplyr::select(trait_j, confounder_trait, trait_j_couple_corr, trait_j_couple_corr_se, 
                corr_due_to_confounding, corr_due_to_confounding_se) %>% 
  mutate(trait_ratio = corr_due_to_confounding/trait_j_couple_corr)

corr_impact_by_traits_ratio <- replace_Neale_ID(corr_impact_by_traits_ratio, traits_corr2_filled, "confounder_trait", "confounder_trait_description") %>% ungroup() %>%
  dplyr::select(-confounder_trait) %>% dplyr::rename(confounder_trait = confounder_trait_description)

corr_impact_ratios <- rbind(corr_impact_by_PCs_ratio, corr_impact_by_coords_ratio, corr_impact_by_traits_ratio)

corr_impact_ratios_summary <- corr_impact_ratios %>%
  group_by(confounder_trait) %>%
  dplyr::summarise(mean_ratio = mean(trait_ratio), median_ratio = median(trait_ratio), sd_ratio = sd(trait_ratio), number_of_traits = n())

corr_impact_ratios_summary %>%
  kbl() %>%
  kable_styling()

# write.table(corr_impact_ratios_summary, "docs/manuscript_tables/corr_impact_ratios_summary.txt", row.names = F, sep=",", quote = F)
```

*A note on the number of traits:*

We tested the impact of `trait_interest` on couple correlation amongst all traits in our couple MR framework (number of traits = `r dim(household_MR_summary_AM)[1]`) with the exception of home location and birth place coordinates (2 each [N and E] * 2), we also removed the trait itself from the confounder analysis where applicable (in the case of traits `20016_irnt`, `738` and `845`). [Trait `189_irnt` was not included in our final couple MR analysis but could still be evaluated here because no MR-estimates were used.]

## Plots.

The figure below shows the 4 traits which showed the greatest impact on couple correlation (mean confounding ratio > 0.01).

```{r confounder_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}

geography_plots <- list()

confounder_traits_sub <- confounder_traits[c(2,3,4,5)]

plots <- list()
for(i in 1:length(confounder_traits_sub)){

  trait <- confounder_traits_sub[i]
  trait_description <- confounder_trait_tibble_descript[which(confounder_trait_tibble_descript[["value"]]==trait),"outcome_description"][[1]]

  if(trait=="738"){trait_description <- "Average total household \nincome before tax"}
  if(trait=="189_irnt"){trait_description <- "Townsend deprivation index \nat recruitment"}
  if(trait=="6160_1"){trait_description <- "Sports club or gym user"}

  plot_data <- corr_impact_by_traits %>% filter(trait_interest==trait) %>% filter(trait_interest!=trait_j)
  plot_data <- replace_Neale_ID(plot_data, traits_corr2_filled, "trait_j", "Phenotype")
  plot <-
    ggplot2::ggplot(
      data = plot_data,
      ggplot2::aes(x = trait_j_couple_corr,
                   y = corr_due_to_confounding, Phenotype = Phenotype)
    ) +
    theme(plot.title = element_text(hjust = 0.5)) +
    geom_point(alpha = 3 / 4) +

    ggplot2::geom_errorbar(
      ggplot2::aes(ymin = corr_due_to_confounding - corr_due_to_confounding_se, ymax = corr_due_to_confounding + corr_due_to_confounding_se),
      colour = "grey",
      width = 0
    ) +
    ggplot2::geom_errorbar(
      ggplot2::aes(xmin = trait_j_couple_corr - trait_j_couple_corr_se, xmax = trait_j_couple_corr + trait_j_couple_corr_se),
      colour = "grey",
      width = 0
    ) +

    theme_minimal() +
    ggplot2::labs(
      x = paste0("Couple trait correlation"),
      y = paste0("Couple correlation \ndue to confounding"),
      title = paste0(trait_description)
      #title = paste0("Impact of ", trait_description, "\n on couple correlation")
    ) +
    xlim(0, 1.0) +
    theme(plot.title = element_text(size=11, hjust = 0.5)) + theme(axis.title = element_text(size = 10))

  plots[[i]] <- plot 
}

i <- 5

figures_grid <- plot_grid(plotlist = plots, nrow=2, ncol = 2, byrow = F, labels="AUTO")
figures_grid
```

```{r PC_confounder_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE}

plot_data <- corr_impact_by_PCs %>% dplyr::select(Neale_pheno_ID, trait_couple_corr, corr_due_to_confounding_all, corr_due_to_confounding_all_se, trait_couple_corr_se, trait_couple_corr_se) %>% unique()

plot_PCs <-
  ggplot2::ggplot(
    data = plot_data,
    ggplot2::aes(x = trait_couple_corr,
                 y = corr_due_to_confounding_all)
  ) +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_point(alpha = 3 / 4) +

  ggplot2::geom_errorbar(
    ggplot2::aes(ymin = corr_due_to_confounding_all - corr_due_to_confounding_all_se, ymax = corr_due_to_confounding_all + corr_due_to_confounding_all_se),
    colour = "grey",
    width = 0
  ) +
  ggplot2::geom_errorbar(
    ggplot2::aes(xmin = trait_couple_corr - trait_couple_corr_se, xmax = trait_couple_corr + trait_couple_corr_se),
    colour = "grey",
    width = 0
  ) +

  theme_minimal() +
  ggplot2::labs(
    x = paste0("Couple trait correlation"),
    y = paste0("Couple correlation \ndue to confounding"),
    title = paste0("Principal components")
    #title = paste0("Impact of ", trait_description, "\n on couple correlation")
  ) +

  xlim(0, 1.0) +
  theme(plot.title = element_text(size=11, hjust = 0.5)) + theme(axis.title = element_text(size = 10))

i <- 5
plots[[i]] <- plot_PCs

```

```{r coord_confounder_plot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE, out.width = "200%", dpi=600}

plot_data <- corr_impact_by_coords %>% dplyr::select(Neale_pheno_ID, trait_couple_corr, corr_due_to_confounding_all, corr_due_to_confounding_all_se, trait_couple_corr_se, trait_couple_corr_se) %>% unique()

plot_coords <-
  ggplot2::ggplot(
    data = plot_data,
    ggplot2::aes(x = trait_couple_corr,
                 y = corr_due_to_confounding_all)
  ) +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_point(alpha = 3 / 4) +

  ggplot2::geom_errorbar(
    ggplot2::aes(ymin = corr_due_to_confounding_all - corr_due_to_confounding_all_se, ymax = corr_due_to_confounding_all + corr_due_to_confounding_all_se),
    colour = "grey",
    width = 0
  ) +
  ggplot2::geom_errorbar(
    ggplot2::aes(xmin = trait_couple_corr - trait_couple_corr_se, xmax = trait_couple_corr + trait_couple_corr_se),
    colour = "grey",
    width = 0
  ) +

  theme_minimal() +
  ggplot2::labs(
    x = paste0("Couple trait correlation"),
    y = paste0("Couple correlation \ndue to confounding"),
    title = paste0("Place of birth \ncoordinates")
    #title = paste0("Impact of ", trait_description, "\n on couple correlation")
  ) +
  xlim(0, 1.0) +
  theme(plot.title = element_text(size=11, hjust = 0.5)) + theme(axis.title = element_text(size = 10))

i <- 6
plots[[i]] <- plot_coords

figures_grid <- plot_grid(plotlist = plots, nrow=3, ncol = 2, byrow = T, labels="AUTO", label_size = 12)
figures_grid

```

# Comparison of paths from index to partner.

As a recap, the model is below.

```{r YiYp_adj_model, echo = FALSE}
library(knitr)
knitr::include_graphics("assets/two_trait_model_yiyp_adj.png", error = FALSE)
```

We sought to compare the different paths from $X_i$ to $Y_p$ using the estimates computed as $\omega$, $\rho$ and $\gamma$.

```{r proxyMR_figure2, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), out.width = "200%", dpi=600}
proxyMR_prod_comparison_fig_yiyp_adj
```

We considered the regression of the following estimates vs $\omega$ with an intercept passing through the origin (i.e. $estimate \sim \omega + 0$:

1. $\rho$
2. $\gamma$
3. $\rho$ + $\gamma$
4. $\rho_{resid} + \gamma$, where $\rho$  is residualized for $\gamma$ (i.e. $\rho_{resid} = resid(lm(\rho \sim \gamma))$)

The results are summarized below:

```{r regression_summary, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

data <- proxyMR_comparison_summary_yiyp_adj_SNPmeta
data$rho_resid <- resid(lm(rho_beta ~ gam_beta, data = data))
data$gam_rho_resid <- data$rho_resid + data$gam_beta

result <- numeric()
for(outcome in c("rho", "gam", "gam_rho", "gam_rho_resid")){

  col_to_rename <- paste0(outcome, "_beta")
  if(outcome=="gam_rho_resid") {col_to_rename <- outcome}
  data_model <- data %>% dplyr::rename(outcome = !!col_to_rename)
  model <- summary(lm(outcome ~ omega_beta + 0, data = data_model)) 
  r2 <- model %>% glance %>% pull(r.squared)
  model_summary <- model %>% tidy() %>% 
    mutate(term = "omega") %>% mutate(outcome = outcome) %>% mutate(r.squared = r2)

  result <- rbind(result, model_summary)

}

omega_regression_estimates <- result %>% dplyr::select(outcome, everything())

omega_regression_estimates %>%
  kbl() %>%
  kable_styling()

```

```{r remove_opp_sign, include = FALSE}

opp_sign_rows <- numeric()
for(col1 in c("gam_beta", "rho_beta", "gam_rho_resid", "omega_beta")){

  for(col2 in c("gam_beta", "rho_beta", "gam_rho_resid", "omega_beta")){

      temp <- which(sign(data[[col1]])!=sign(data[[col2]]))
      opp_sign_rows <- c(opp_sign_rows, temp)
  }

}

opp_sign_rows <- unique(opp_sign_rows)

```

To further explore these estimates, performed paired t-tests to find the significant differences between pairs of estimates. We first removed all instances where the sign did not match between any two pairings. In total, of the `r dim(data)[1]` trait pairs under consideration, `r length(opp_sign_rows)` were removed.

```{r boxplot, echo = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), message=FALSE, warning = FALSE}

df <- data[-opp_sign_rows,] %>% dplyr::select(exposure_ID, outcome_ID, gam_beta, rho_beta, gam_rho_resid, omega_beta)

data_boxplot  <- df %>%
  pivot_longer(cols = !c("exposure_ID", "outcome_ID"), names_to = "data", values_to = "beta") %>%
  mutate_at("data", as.factor) %>%
  mutate(
    data = fct_recode(as.factor(data),
                      Gamma = "gam_beta",
                      Rho = "rho_beta",
                      Gamma_Rho_resid = "gam_rho_resid",
                      Omega = "omega_beta")
  )

pairwise.t.test(data_boxplot$beta, data_boxplot$data, p.adjust = "none")

#cmpr <- list(c("Omega", "Gamma_Rho"), c("Gamma","Gamma_Rho"), c("Gamma_Rho","Gamma_Rho_resid"), c("Rho", "Gamma_Rho"))

position <- data_boxplot %>% group_by(data) %>% dplyr::summarise(median=median(beta)) %>% 
  arrange(-desc(median)) %>% pull(data)

data_boxplot$data <- factor(data_boxplot$data,levels=position)

data_boxplot <- data_boxplot %>% filter(data != "Gamma_Rho")

boxplot <- data_boxplot %>%
  ggplot( aes(x=data, y=beta)) +
  geom_boxplot(outlier.shape = NA) +
  coord_cartesian(ylim=c(-0.5, 0.5)) +
  scale_fill_viridis(discrete = TRUE, alpha=0.6, option="A") +
  theme_ipsum() +
  theme(
    legend.position="none",
    plot.title = element_text(size=11)
  ) +
  #labs(x ="Estimate", y = "") +
  scale_x_discrete(labels=c("Omega" = "\u03C9", 
                            "Gamma" = "\u03B3", "Gamma_Rho_resid" = expression( paste(rho ["resid"], " + ", gamma)), "Rho" = "\u03C1")) +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank())

#+ #"Gamma_Rho" = "\u03B3 + \u03C1",

  # stat_compare_means(comparisons = cmpr, tip.length=0.01,
  #                    label = "p.signif",
  #                    symnum.args = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1),
  #                                       symbols = c("****", "***", "**", "*", "ns"))) 
  #

boxplot

figures_grid <- plot_grid(plotlist = list(proxyMR_prod_comparison_fig_yiyp_adj, boxplot), nrow=1, ncol = 2, byrow = F)
figures_grid

```

## Paths summary.

A summary of the $\omega$, $\rho$, $\gamma$ and $\rho_{resid} + \gamma$ estimates are summarized in the table below. The differences between the estimates (according to a Z-test) are given in the last three columns. Trait pairs were pruned such that:  if there is a pair A-B and C-D and if [`max(corr(A,C)*corr(B,D),corr(A,D)*corr(B,C))`] > `r corr_trait_threshold` then one pair is dropped. Pairs were prioritized based on the $\omega$ p-value.

```{r paths_summary, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}

tar_load(proxyMR_comparison_summary_yiyp_adj_SNPmeta_prune)

data <- proxyMR_comparison_summary_yiyp_adj_SNPmeta_prune
data$rho_resid <- resid(lm(rho_beta ~ gam_beta, data = data))
data$gam_rho_resid <- data$rho_resid + data$gam_beta

paths_summary <- data %>% dplyr::select(exposure_description, outcome_description, gam_beta, rho_beta, omega_beta, gam_rho_resid, gam_vs_rho_pval, omega_vs_rho_pval, omega_vs_gam_pval, omega_vs_gam_rho_resid_pval)

format_round_cols <- c(colnames(dplyr::select(paths_summary, ends_with("_beta"))), "gam_rho_resid")
format_signif_cols <- colnames(dplyr::select(paths_summary, ends_with("_pval")))

DT::datatable(paths_summary, width = "100%") %>% formatRound(columns=format_round_cols, digits = 3) %>%
  formatSignif(columns = format_signif_cols, digits = 3)

```

```{r write_tables, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}

################################# 
## Tables to include in paper ###
#################################

## Cases where correlation is > MR effects: `corr_vs_MR_effects_sig_dat`

## Potential confounders for all cases where corr is > MR effects: `sig_confounders`

## Significant same-trait couple MR effects: `AM_result_sig.DT`

## Nominally significant same-trait sex-difference: `nominally_sig_sex_differences`

## Phenotypic correlations by bin: `binned_pheno_corr_sig.DT`

## Ratio of confounding to correlation: `corr_impact_ratios_summary`

## Omega regression estimates: `omega_regression_estimates`

## Pruend paths summary: `paths_summary`

file_name <- "docs/manuscript_tables/manuscript_tables.xlsx"

write.xlsx_wrapper <- function(dat, ...){

  write.xlsx(as.data.frame(dat), ...)
}

write.xlsx_wrapper(corr_vs_MR_effects_sig_dat, file=file_name, sheetName="sig_corr_vs_MR", row.names=FALSE)
write.xlsx_wrapper(sig_confounders, file=file_name, sheetName="sig_confounders", append=TRUE, row.names=FALSE)
write.xlsx_wrapper(AM_result_sig.DT, file=file_name, sheetName="BF_sig_same_trait_MR", append=TRUE, row.names=FALSE)
write.xlsx_wrapper(nominally_sig_sex_differences, file=file_name, sheetName="nominally_sig_sex_diff", append=TRUE, row.names=FALSE)
write.xlsx_wrapper(binned_pheno_corr_sig.DT, file=file_name, sheetName="pheno_binned_corrs", append=TRUE, row.names=FALSE)
write.xlsx_wrapper(corr_impact_ratios_summary, file=file_name, sheetName="corr_impact_ratios_summary", append=TRUE, row.names=FALSE)
write.xlsx_wrapper(omega_regression_estimates, file=file_name, sheetName="omega_regression_estimates", append=TRUE, row.names=FALSE)
write.xlsx_wrapper(paths_summary, file=file_name, sheetName="paths_summary", append=TRUE, row.names=FALSE)

```