### 2022-03-07 analysis/index.Rmd
wflow_publish("analysis/index.Rmd")

---
title: "Home"
author: "Jenny Sjaarda"
site: workflowr::wflow_site
output:
  bookdown::html_document2:
    toc: true
editor_options:
  chunk_output_type: console
---

This is the website for the research project "proxyMR".

If you have cloned the project to a local computer this website is rendered in the `docs` subdirectory of the project directory.

If you are using `workflowr` to publish the research website it will also be rendered online to GitHub Pages.

This page acts as a table of contents for the website. There are links to the web pages generated from the analysis notebooks and to the rendered versions of manuscripts/documents/presentations.

------------------------------------------------------------------------

# Overview documents {.unnumbered}

## [About](about.html) {.unnumbered}

This notebook explains the central ideas behind the project.

## [Notes](notes.html) {.unnumbered}

This notebook is for keeping notes of any points that may be useful for
later project or manuscript development.

## [To do](todo_list.html) {.unnumbered}

Current tasks and completed tasks (roughly organized by dates/weeks).

## [Project Workflow Status](project_status.html) {.unnumbered}

This notebook displays the computational status of the project, that is, whether everything is up to date. Ironically, this is currently the only notebook that must be run manually, so this notebook only displays the status **when the notebook was last executed** and there is no indication whether the project status has changed since then.

This project uses the [`targets`](https://wlandau.github.io/targets/) and [`workflowr`](https://github.com/jdblischak/workflowr) packages for managing the workflow of the project (making sure that the dependencies between computational steps are satisfied).

## [Meetings](meetings.html) {.unnumbered}

Weekly meeting were held roughly once per week or every other week.

## [Manuscript](manuscript.html) {.unnumbered}

Document with a final summary of the analyses and tables for the manuscript.

------------------------------------------------------------------------

# Publications and Presentations {.unnumbered}

- Group Meeting [07-09-2021](presentations/group_meeting_070921/group_meeting_proxyMR_070921.pdf)

------------------------------------------------------------------------

# Analyses Notebooks {.unnumbered}

These notebooks capture the analyses that were carried out to develop
the code of the core [processing pipeline](pipeline_overview.html).

## 1. Prepare data {.unnumbered}

Prepare phenotypes and couples for analysis in MR pipeline.

### [Couple selection](couple_selection.html) {.unnumbered}

Identify couples in the UKBB.

### [Phenotype selection](phenotype_selection.html) {.unnumbered}

Select phenotypes for analysis in pipeline.

### [Final phenotype list](final_phenotype_list.html) {.unnumbered}

Final list of phenotypes for analysis in pipeline and some basic characteristics (correlation among couples, number of independent IVs).

## 2. Run MR pipeline {.unnumbered}

### [MR between partners (partner vs index)]() {.unnumbered}

- Identify SNP-exposure effects (GW-significance) for phenotypes (sex specific) from Neale database [$X_{i} \sim G_{i}$].
- Test SNP-outcome effects between genetic instruments and partner outcome phenotypes (sex specific) [$Y_{p} \sim G_{i}$].
- Combine in MR to assess causal relationships between index and partner [$Y_{p} \sim X_{i}$].

### [Assortative mating MR](AM_MR_summary.html) {.unnumbered}

- Restrict to only MRs where exposure is the same as outcome (for downstream analyses) [$X_{p} \sim X_{i}$].

### [MR within same individual]() {.unnumbered}

- Identify SNP-exposure effects (GW-significance) for phenotypes (sex specific) from Neale database [$X_{i} \sim G_{i}$].
- Pull SNP-outcome effects  between genetic instruments and outcome phenotypes (sex specific) from Neale database [$Y_{i} \sim G_{i}$].
- Combine in MR to assess causal relationship between all phenotypes [$Y_{i} \sim X_{i}$].

### [Compare paths towards outcome in partner]() {.unnumbered}

- Compare paths from index exposure phenotype to partner outcome phenotype [$X_{i} \rightarrow Y_{p}$].

### [Assess impact of geography on causal relationships between couples](geographical_impact.html) {.unnumbered}

------------------------------------------------------------------------

# Acknowledgements {.unnumbered}

Primary analysis performed by: [Jenny Sjaarda](https://wp.unil.ch/sgg/people/jenny-sjaarda/).

Project supervisor: [Professor Zoltan Kutalik](https://wp.unil.ch/sgg/people/zoltan/).

This work has been supported by the [Swiss National Science Foundation](http://www.snf.ch/en/funding/Pages/default.aspx).