### 2022-04-20 analysis/project_status.Rmd

---
title: "Project status"
author: "Jenny Sjaarda"
date: "2021-09-16"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

This notebook displays the status of the project workflow, that is, whether everything is computationally up to date. It displays the status as at the time the notebook is executed, so if the status has changed since this notebook was last executed that fact won't be reflected in this rendered notebook. This notebook is currently intended to be run manually when you feel the need, for example, when returning to the project.

# `{renv}` status

`{renv}` tracks the used `R` packages and their versions.
```{r renv}
# Get {targets} to tell {renv} which packages it uses
targets::tar_renv() 
renv::status()
```

# `{targets}` status
`{targets}` is used to track the computational dependencies between the
project components and orchestrate recalculation when any components are
changed.

Each of the meta pipelines has leaves that are reports rendered from
`Rmd` files that are managed jointly by `{targets}` and `{workflowr}`.
`{targets}` is aware of the upstream dependencies (data and functions)
and the `Rmd` file itself. If any of these is invalidated `tar_make()`
will rebuild the report using `wflow_build()`
.
`{targets}` is *not* aware of the rendered meta report files. The
rendered report files are managed exclusively by `{workflowr}`. This
means that you can use `{workflowr}` to build and publish the report
without making `{targets}` think that the report is out of date because
`{workflowr}` has re-rendered it.
```{r targets}
# outline
targets::tar_glimpse()
# excruciating detail
targets::tar_visnetwork()
```

# `{workflowr}` status
`{workflowr}` is used to render the *meta* documents
(`analysis/m_*.Rmd`) and build the project website.
`{workflowr}` is only concerned with the modification and `git` commit
dates of the `Rmd` and `HTML` files, whereas `{targets}` is concerned
with the computational dependencies. Thus, you should rely on the
`{targets}` status to ensure the meta reports are computationally up to
date, then rely on the `{workflowr}` status to manage the building and
publication of the entire website.

```{r workflowr}
workflowr::wflow_status()
```