### 2022-04-20 analysis/AM_MR_summary.Rmd
wflow_publish("analysis/AM_MR_summary.Rmd")

---
title: "Assortative mating MR summary"
author: "Jenny Sjaarda"
date: "2021-09-15"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}

options(scipen=999)
source("code/settings.R")
source("R/functions.R")
library(targets)
library(dplyr)
library(DT)
library(knitr)

```

```{r load_targets, include = FALSE, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE)}

tar_load(traits_final)
tar_load(household_MR_binned_het)
tar_load(household_MR_summary_AM)
tar_load(traits_corr2_filled)
tar_load(num_tests_by_PCs)
tar_load(num_tests_by_PCs_AM_sig)
tar_load(exposures_to_run)
tar_load(household_MR_AM_FvsM_fig)
tar_load(household_MR_binned_het_sex_spec)

```

# Background

For certain analyses it is worth only looking at MR results between couples where the exposure trait is the same as the outcome trait. We've called these analyses *assortative mating MR (AM MR)*.

```{r single_trait_fig, echo = FALSE}
library(knitr)
knitr::include_graphics("assets/single_trait_model.png", error = FALSE)
```

# Results

```{r AM_data_prep, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE, include = FALSE}

AM_sig_traits <- household_MR_summary_AM %>% filter(IVW_pval < 0.05/num_tests_by_PCs) %>% pull(exposure_ID) %>% unique()
AM_result_sig <- household_MR_summary_AM %>% filter(exposure_ID %in% AM_sig_traits)

household_MR_binned_het_temp <- replace_Neale_ID(household_MR_binned_het, traits_corr2_filled, "exposure_ID", "exposure_description")
household_MR_binned_het_temp2 <- replace_Neale_ID(household_MR_binned_het_temp, traits_corr2_filled, "outcome_ID", "outcome_description")
household_MR_binned_het_descript <- household_MR_binned_het_temp2

AM_sex_spec_result_sig <-
  household_MR_binned_het_sex_spec %>% filter(same_trait == TRUE) %>% filter(exposure_ID %in% AM_sig_traits) %>% filter(sex_het_pval < 0.05 /
                                                                                                                          num_tests_by_PCs_AM_sig)

```

- The number of analyses corresponds to the number of traits (`r dim(traits_final)[1]`). 
- The significant results after adjusting for multiple hypothesis testing are shown below. 
- `IVW_meta_beta` and `IVW_meta_pval` corresponds to the beta and p-value of the meta-analyzed MR across sexes, respectively (i.e. MR estimates were computed in each sex-seperately using sex-specific SNP-exposure and SNP-outcome results and then meta-analyzed).  
- After adjusting for multiple tests (p < `0.05/`r num_tests_by_PCs``), identified `r length(AM_sig_traits)` significant AM MR relationships (see below on details for how number of effective tests was obtained).

```{r AM_overview, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}

AM_result_sig.DT <- AM_result_sig %>% dplyr::select(exposure_description, IVW_beta, IVW_pval) %>% unique()
DT::datatable(AM_result_sig.DT, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval'), digits=3)

AM_sig_sex_het_nom <- household_MR_binned_het_sex_spec %>% filter(same_trait) %>% filter(exposure_ID %in% AM_sig_traits) %>% filter(sex_het_pval < 0.05)
AM_sig_sex_het_nom <- replace_Neale_ID(AM_sig_sex_het_nom, traits_corr2_filled,"exposure_ID", "exposure_description") %>%
  dplyr::rename(trait_description = exposure_description) %>%
  dplyr::rename(trait_ID = exposure_ID) %>%
  dplyr::select(trait_description, trait_ID, everything())

AM_result_sig_nom.DT <- AM_sig_sex_het_nom %>% dplyr::select(trait_description, trait_ID, exposure_sex, IVW_beta, IVW_se, IVW_pval, sex_het_pval) %>% unique()

```

**A note on multiple testing adjustment:**

  - Adjustment was based on PCs of correlation matrix.  
  - Calculate correlation matrix of all `r dim(traits_final)[1]` traits tested. 
  - Perform PCs on the resulting corr matrix. 
  - Number of tests = # of PCs needed until variance explained is > 99.5%. 
  - Resulting # of effective tests: `r num_tests_by_PCs`.

## Sex heterogeneity

- For each AM MR ($X_{p} \sim X_i$), analyses were run in each sex separately. 
- Of the `r length(AM_sig_traits)` significant results shown above, we tested to see if there was any difference in AM MR estimates between sexes. 
- A similar adjustment for multiple hypothesis testing was applied as above, resulting in `r num_tests_by_PCs_AM_sig` number of effective tests. 
- After adjusting for multiple hypothesis testing (p < `0.05/`r num_tests_by_PCs_AM_sig``), `r dim(AM_sex_spec_result_sig)[1]` traits showed significant differences amongst sexes.
- The table below shows the *nominally significant* results at `p < 0.05` (of which there are `r dim(AM_result_sig_nom.DT)[1]/2`).

```{r AM_sex_het_nom, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}

DT::datatable(AM_result_sig_nom.DT, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval', 'sex_het_pval'), digits=3)

nominally_sig_sex_differences <- AM_result_sig_nom.DT %>% pivot_wider(names_from = exposure_sex, values_from = c(IVW_beta, IVW_se, IVW_pval))

t.test(nominally_sig_sex_differences$IVW_beta_male, nominally_sig_sex_differences$IVW_beta_female, paired = T)

```

In general, found that female estimates are on average larger than males (result of the paired t-test shown below).

```{r sex_het_m_vs_f, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}

nominally_sig_sex_differences <- AM_result_sig_nom.DT %>% pivot_wider(names_from = exposure_sex, values_from = c(IVW_beta, IVW_se, IVW_pval))

t.test(nominally_sig_sex_differences$IVW_beta_male, nominally_sig_sex_differences$IVW_beta_female, paired = T)

```

## Binned results

- For each AM MR ($X_{p} \sim X_i$), analyses were run in the full sample as well as in 5 roughly equal sized bins according to time couple had been together (`time_together_even_bins`) estimated using [time at household variable](https://biobank.ctsu.ox.ac.uk/showcase/field.cgi?id=699), and median age (`age_even_bins`). 
- Of the `r length(AM_sig_traits)` significant results shown above, we tested to see if there was any significant difference in AM MR estimates amongst the two grouping variables. 
- Specifically, the outcome data set was split into 5 bins, and the SNP-outcome effect was estimated in each bin separately (and each sex seperately). 
- These outcome-SNP effects were then used to generate bin-specific MR estimates, using the same SNP-exposure effects from Neale. 
- To estimate the difference among bins, we used two approaches: 
    1. The Cochran's Q test to test for heterogeneity in meta-analyses.
    2. Tested the significance of the slope of linear model of median bin (either `age` or `time together`) versus the bin-specific MR estimate.  
- The two results are shown below.

### Q-statistic across bins

```{r bin_q_het_setup, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}

q_het_result_age <- household_MR_binned_het_descript %>% filter(same_trait==TRUE) %>% filter(grouping_var=="age_even_bins") %>% filter(IVW_meta_pval < 0.05/num_tests_by_PCs) %>%
  dplyr::select(exposure_description, exposure_sex, outcome_sex, grouping_var, IVW_beta, IVW_pval, Q_pval, Q_pval_meta) %>% arrange(Q_pval_meta)

q_het_result_tt <- household_MR_binned_het_descript %>% filter(same_trait==TRUE) %>% filter(grouping_var=="time_together_even_bins") %>% filter(IVW_meta_pval < 0.05/num_tests_by_PCs) %>%
  dplyr::select(exposure_description, exposure_sex, outcome_sex, grouping_var, IVW_beta, IVW_pval, Q_pval, Q_pval_meta) %>% arrange(Q_pval_meta)

```

The number of significant q-statistics significant after multiple hypothesis testing (p < `0.05/`r num_tests_by_PCs_AM_sig` = `r 0.05/num_tests_by_PCs_AM_sig``) in each group was:

  - Binned by median *age* of couples, number of significant results within meta = `r length(which(q_het_result_age$Q_pval_meta < 0.05/num_tests_by_PCs_AM_sig))/2` and sex-specific = `r length(which(q_het_result_age$Q_pval < 0.05/num_tests_by_PCs_AM_sig))`.
  - Binned by *time* couple had been together, number of significant results within meta = `r length(which(q_het_result_tt$Q_pval_meta < 0.05/num_tests_by_PCs_AM_sig))/2` and sex-specific = `r length(which(q_het_result_tt$Q_pval < 0.05/num_tests_by_PCs_AM_sig))`.

All results are displayed below (sort by Q_pval to view significant results).

```{r bin_q_het, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}

DT::datatable(q_het_result_age, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval', 'Q_pval', 'Q_pval_meta'), digits=3)
DT::datatable(q_het_result_tt, width = "100%") %>% formatSignif(columns=c('IVW_beta', 'IVW_pval', 'Q_pval', 'Q_pval_meta'), digits=3)
```

### Trend across bins

```{r bin_slope_setup, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}

slope_result_age <- household_MR_binned_het_descript %>% filter(same_trait==TRUE) %>% filter(grouping_var=="age_even_bins") %>% filter(IVW_meta_pval < 0.05/num_tests_by_PCs) %>%
  dplyr::select(exposure_description, exposure_sex, grouping_var, bin_slope_beta, bin_slope_pval, bin_slope_meta_pval, bin_slope_meta_pval_wt, bin_slope_sex_het_pval) %>% arrange(bin_slope_meta_pval)

slope_result_tt <- household_MR_binned_het_descript %>% filter(same_trait==TRUE) %>% filter(grouping_var=="time_together_even_bins") %>% filter(IVW_meta_pval < 0.05/num_tests_by_PCs) %>%
  dplyr::select(exposure_description, exposure_sex, grouping_var, bin_slope_beta, bin_slope_pval, bin_slope_meta_pval, bin_slope_meta_pval_wt, bin_slope_sex_het_pval) %>% arrange(bin_slope_meta_pval)

```

Slopes were calculated by estimating the beta-coefficient of a linear regression between MR estimate within each bin (dependent variable) and median bin (independent variable, either age or time at same household). Linear models were run both *unweighted* and *weighted* for the by the inverse of the SE of the MR estimate.

The number of significant trends ($\beta$ estimates from the model: $\alpha_{bin} \sim median_{bin}$) significant after multiple hypothesis testing (p < `0.05/`r num_tests_by_PCs_AM_sig` = `r 0.05/num_tests_by_PCs_AM_sig``) in each group was:

  - Binned by median *age* of couples, number of significant results after adjusting for number of tests:
      - $\beta$ unweighted, meta-anlayzed across sexes: `r length(which(slope_result_age$bin_slope_meta_pval < 0.05/num_tests_by_PCs_AM_sig))/2`
      - $\beta$ weighted by inverse of the SE, meta-anlayzed across sexes: `r length(which(slope_result_age$bin_slope_meta_pval_wt < 0.05/num_tests_by_PCs_AM_sig))/2`
      - $\beta$ significantly different between sexes: `r length(which(slope_result_age$bin_slope_sex_het_pval < 0.05/num_tests_by_PCs_AM_sig))/2`
  - Binned by *time* couple had been together, number of significant results after adjusting for number of tests:
      - $\beta$ unweighted, meta-anlayzed across sexes: `r length(which(slope_result_tt$bin_slope_meta_pval < 0.05/num_tests_by_PCs_AM_sig))/2`
      - $\beta$ weighted by inverse of the SE, meta-anlayzed across sexes: `r length(which(slope_result_tt$bin_slope_meta_pval_wt < 0.05/num_tests_by_PCs_AM_sig))/2`
      - $\beta$ significantly different between sexes: `r length(which(slope_result_tt$bin_slope_sex_het_pval < 0.05/num_tests_by_PCs_AM_sig))/2`.

All results are displayed below (sort by various p-values to view significant results).

```{r bin_slope, eval = ifelse(Sys.info()["sysname"]=="Linux", TRUE, FALSE), echo = FALSE}

DT::datatable(slope_result_age, width = "100%") %>% formatSignif(columns=c('bin_slope_beta', 'bin_slope_pval', 'bin_slope_meta_pval', 'bin_slope_meta_pval_wt', 'bin_slope_sex_het_pval'), digits=3)
DT::datatable(slope_result_tt, width = "100%") %>% formatSignif(columns=c('bin_slope_beta', 'bin_slope_pval', 'bin_slope_meta_pval', 'bin_slope_meta_pval_wt', 'bin_slope_sex_het_pval'), digits=3)
```

### Figures

A few figures of interest are shown below.

```{r binned_figures, include = T, echo = FALSE}

target1 <- which(exposures_to_run$Neale_pheno_ID=="1080")
figs1 <- tar_read(household_MR_binned_AM_figs, branches = target1)[[1]]

figs1[[5]]
figs1[[7]]

target2 <- which(exposures_to_run$Neale_pheno_ID=="1239")
figs2 <- tar_read(household_MR_binned_AM_figs, branches = target2)[[1]]

figs2[[5]]
figs2[[7]]

target3 <- which(exposures_to_run$Neale_pheno_ID=="20153_irnt")
figs3 <- tar_read(household_MR_binned_AM_figs, branches = target3)[[1]]
figs3[[5]]
figs3[[9]]

target4 <- which(exposures_to_run$Neale_pheno_ID=="1050")
figs4 <- tar_read(household_MR_binned_AM_figs, branches = target4)[[1]]
figs4[[4]]
figs4[[8]]
```